---
title: "Candidate file prep for EPTS inequity"
output:
  html_document:
    toc: true
    df_print: paged
authors: Nour Asfour, Kevin Zhang, and William Parker
---


```{r package_load}
library(tidyverse)
library(lubridate)
library(haven)
```

```{r key_parameters_patient_population}
start_date <- '2015-01-01'
end_date <- '2020-12-31'
living_donor_recips <- TRUE
```


# Identify Adult Kidney-alone registrations from `r start_date` to `r end_date`
```{r load_in_data}
df_cand_kipa_all <- haven::read_sas('./cand_kipa.sas7bdat') %>%
  filter(as.Date(CAN_LISTING_DT) >= start_date & as.Date(CAN_LISTING_DT) <= end_date) %>%
  filter(WL_ORG == 'KI') %>%
  filter(CAN_AGE_AT_LISTING >= 18)

num_adult_registrations <- df_cand_kipa_all$PX_ID %>% unique() %>% length()
```


# create waitlist time end date for each `PX_ID` registration
```{r}
df_cand_kipa <- df_cand_kipa_all %>%
  mutate(waitlist_end_date = case_when(
    is.na(REC_TX_DT) == FALSE ~ REC_TX_DT,
    is.na(CAN_REM_DT) == FALSE ~ CAN_REM_DT,
    is.na(CAN_LAST_INACT_STAT_DT) == FALSE & CAN_LAST_INACT_STAT_DT > CAN_LAST_ACT_STAT_DT ~ CAN_LAST_INACT_STAT_DT,
    TRUE ~CAN_LAST_ACT_STAT_DT
  ))
```



# create wait time and outcome variable for candidates with a single registration
```{r}
single_registrations <- df_cand_kipa %>%
  group_by(PERS_ID) %>%
  mutate(num_list = n()) %>%
  filter(num_list == 1) %>%
  ungroup() %>% 
  mutate(min_list_date = CAN_LISTING_DT,
         wait_time = waitlist_end_date - min_list_date,
         outcome = case_when(
           DON_TY == "C" ~ "DDKT",
           DON_TY == "L" ~ "LDKT",
           is.na(CAN_REM_CD) == FALSE ~ "removed/died",
           TRUE ~ "censored"
         ))

single_registrations %>%
  select(CAN_LISTING_DT, min_list_date, CAN_REM_DT, CAN_REM_CD, REC_TX_DT, waitlist_end_date, wait_time, outcome) %>%
  count(outcome)
```



# Create a single record for candidates with multiple concurrent registrations
```{r}
multiple_registrations <- df_cand_kipa %>%
  filter(!PX_ID %in% single_registrations$PX_ID) %>%
  group_by(PERS_ID) %>%
  arrange(PERS_ID, CAN_LISTING_DT) %>%
  mutate(num_list = n())


candidates_w_multiple_registrations <- multiple_registrations %>% pull(PERS_ID) %>% unique() %>% length()

avg_num_registrations_multiple <- mean(multiple_registrations %>% group_by(PERS_ID) %>% filter(row_number() ==1) %>% pull(num_list))
```


## Seperate sequential and consecutive registrations
```{r}
########################################################################
# multiple_registrations_oldsort <- multiple_registrations
# multiple_registrations_oldsort <- 
#   multiple_registrations_oldsort[order(multiple_registrations_oldsort$PERS_ID, multiple_registrations_oldsort$waitlist_end_date), ]
# multiple_registrations_oldsort <- multiple_registrations_oldsort %>%
#   mutate(list_type = case_when(
#     CAN_LISTING_DT < lag(waitlist_end_date) ~ "concurrent",
#     waitlist_end_date > lead(CAN_LISTING_DT) ~ "concurrent",
#     TRUE ~ "sequential")) %>%
#   mutate(REC_TX_DT = as.Date(REC_TX_DT)) %>%
#   mutate(num_tx = length(unique(na.omit(REC_TX_DT)))) %>%
#   fill(REC_TX_DT, .direction='downup')
########################################################################



multiple_registrations <- multiple_registrations %>%
  mutate(list_type = case_when(
    CAN_LISTING_DT < lag(waitlist_end_date) ~ "concurrent",
    waitlist_end_date > lead(CAN_LISTING_DT) ~ "concurrent",
    TRUE ~ "sequential")) %>%
  mutate(REC_TX_DT = as.Date(REC_TX_DT)) %>%
  mutate(num_tx = length(unique(na.omit(REC_TX_DT)))) %>%
  fill(REC_TX_DT, .direction='downup')


## Relocate order (sort) function
multiple_registrations <- multiple_registrations[order(multiple_registrations$PERS_ID, multiple_registrations$waitlist_end_date), ]



## A retransplant counter
multiple_registrations$transplant_num <- 1


## If transplant date changed from previous row to current row but person ID stayed the same, counter + 1
for(i in 2:nrow(multiple_registrations)) {
  if(multiple_registrations$PERS_ID[i-1] == multiple_registrations$PERS_ID[i] &
     multiple_registrations$REC_TX_DT[i-1] != multiple_registrations$REC_TX_DT[i] &
     !is.na(multiple_registrations$REC_TX_DT[i])) {
    
    multiple_registrations$transplant_num[i] = multiple_registrations$transplant_num[i-1] + 1
     } 
}

## Previous step only changed the counter at the first instance where transplant date changed
## So now we fill 'down' this new counter value
## If the last row had a different counter value but person ID stayed the same, change this row's counter to the last row's
for(i in 2:nrow(multiple_registrations)) {
  if(multiple_registrations$PERS_ID[i-1] == multiple_registrations$PERS_ID[i] &
     multiple_registrations$transplant_num[i-1] != multiple_registrations$transplant_num[i] &
     multiple_registrations$transplant_num[i-1] != 1) {
    
    multiple_registrations$transplant_num[i] = multiple_registrations$transplant_num[i-1]
     } 
}

## Change the counter value back to 0 for sequential
multiple_registrations$transplant_num[multiple_registrations$list_type == 'sequential'] <- 0



sequential_lists <- multiple_registrations %>%
  filter(list_type == "sequential") %>%
  mutate(min_list_date = CAN_LISTING_DT,
         wait_time = waitlist_end_date - min_list_date,
         outcome = case_when(
           DON_TY == "C" ~ "DDKT",
           DON_TY == "L" ~ "LDKT",
           is.na(CAN_REM_CD) == FALSE ~ "removed/died",
           TRUE ~ "censored"
         ))


max_retransplants <- max(multiple_registrations$transplant_num)

## Minimum list date for each concurrent transplant
multiple_registrations <- multiple_registrations %>%
  group_by(PERS_ID, transplant_num) %>%
  mutate(min_list_date = min(CAN_LISTING_DT, na.rm=T))


collapsed_concurrent_registrations <- NULL
for(i in 1:max_retransplants) {
  
  collapsed_concurrent_registrations <- rbind(collapsed_concurrent_registrations, 
        
  multiple_registrations %>%
  filter(list_type == "concurrent" & transplant_num == i) %>% ## Do it seperately for each transplant counter number
  mutate(DON_TY = ifelse(DON_TY == "", NA, DON_TY),
         last_wait_date = max(waitlist_end_date, na.rm = TRUE))%>%
  fill(REC_TX_DT, .direction = "up") %>%
  fill(DON_TY, .direction = "up") %>%
  fill(DONOR_ID, .direction = "up") %>%
  fill(CAN_REM_CD, .direction = "up") %>%
  #filter(row_number() ==1) %>%
  mutate(wait_time = case_when(
    is.na(REC_TX_DT) == FALSE & transplant_num != '0' ~ REC_TX_DT- min_list_date, ### Added, ignore sequential rows
    TRUE ~ last_wait_date - min_list_date),
    outcome = case_when(
           DON_TY == "C" ~ "DDKT",
           DON_TY == "L" ~ "LDKT",
           is.na(CAN_REM_CD) == FALSE ~ "removed/died",
           TRUE ~ "censored")
    ) %>%
  select(-c(waitlist_end_date, CAN_LISTING_DT, CAN_REM_DT)) %>%
  filter(row_number() ==1) %>%
  mutate(last_wait_date = case_when(
    REC_TX_DT < last_wait_date ~ REC_TX_DT,
    TRUE ~last_wait_date))) 
  
}

```

## Code check for typical case
```{r}
df_cand_kipa %>%
  filter(PERS_ID == 2235145) %>%
  select(PERS_ID, CAN_LISTING_DT, REC_TX_DT, DON_TY, CAN_REM_CD, CAN_LAST_ACT_STAT_DT, CAN_LAST_INACT_STAT_DT)
```

```{r}
collapsed_concurrent_registrations %>%
  ungroup() %>%
  select(PERS_ID, min_list_date, REC_TX_DT, DON_TY, CAN_REM_CD, last_wait_date, wait_time, outcome) %>%
  filter(PERS_ID == 2235145)
```


## Resolved edge case - fixed
- Candidate with multiple rounds of concurrent listings gets collapsed into a single observation `PERS_ID == 2255750`
```{r}
multiple_registrations %>% 
  group_by(PERS_ID) %>%
  mutate(list_num = row_number(),
         num_tx = n_distinct(REC_TX_DT, na.rm = TRUE)) %>%
  select(PERS_ID, list_num, list_type, CAN_LISTING_DT, waitlist_end_date, num_tx, REC_TX_DT, DONOR_ID, DON_TY) %>%
  filter(PERS_ID == '2255750')
```

### Second concurrent listing is included with correct data
```{r}
collapsed_concurrent_registrations %>%
  select(PERS_ID, min_list_date, last_wait_date, REC_TX_DT, wait_time, outcome, DONOR_ID) %>%
  filter(PERS_ID == 2255750)
```



## Check a second edge case example to confirm 
Candidate where the second waitlist observation ended after the third waitlist observation (before sorting), and the listing date requires a min listing date
```{r}
multiple_registrations %>%
  group_by(PERS_ID) %>%
  arrange(PERS_ID, CAN_LISTING_DT) %>% 
  mutate(list_num = row_number(),
         num_tx = n_distinct(REC_TX_DT, na.rm = TRUE)) %>%
  select(PERS_ID, list_num, list_type, CAN_LISTING_DT, min_list_date, waitlist_end_date, num_tx, REC_TX_DT, DONOR_ID, DON_TY) %>%
  filter(PERS_ID == 2232160)
```


```{r}
collapsed_concurrent_registrations %>%
  select(PERS_ID, min_list_date, last_wait_date, REC_TX_DT, wait_time, outcome, DONOR_ID) %>%
  filter(PERS_ID == 2232160)
```

# Look for other inconsistent cases
```{r}
sample_check <-df_cand_kipa %>%
  select(PERS_ID, CAN_LISTING_CTR_ID, CAN_LISTING_DT, CAN_REM_DT, REC_TX_DT, DON_TY, CAN_REM_CD, CAN_LAST_ACT_STAT_DT, CAN_LAST_INACT_STAT_DT) %>%
    group_by(PERS_ID) %>%
    mutate(total_list = n()) %>%
    filter(total_list > 5)

sample_collapsed_concurrent_PERS_ID <- collapsed_concurrent_registrations %>%
  filter(PERS_ID %in% sample_check$PERS_ID) %>%
  select(PERS_ID, min_list_date, last_wait_date, REC_TX_DT, wait_time, outcome, DONOR_ID) %>%
  ungroup()

sample_check %>% 
  arrange(PERS_ID, CAN_LISTING_DT)
```


## Old dataset with earlier sort- list num 1 and 2 should be concurrent, while 6 should be sequential
```{r}
# multiple_registrations_oldsort %>%
#   group_by(PERS_ID) %>%
#   arrange(PERS_ID, CAN_LISTING_DT) %>% 
#   mutate(list_num = row_number(),
#          num_tx = n_distinct(REC_TX_DT, na.rm = TRUE)) %>%
#   select(PERS_ID, list_num, list_type, CAN_LISTING_DT, waitlist_end_date, num_tx, REC_TX_DT, DONOR_ID, DON_TY) %>%
#   filter(PERS_ID==5117188)
```


## New dataset
```{r}
multiple_registrations %>%
  group_by(PERS_ID) %>%
  arrange(PERS_ID, CAN_LISTING_DT) %>% 
  mutate(list_num = row_number(),
         num_tx = n_distinct(REC_TX_DT, na.rm = TRUE)) %>%
  select(PERS_ID, list_num, list_type, CAN_LISTING_DT, min_list_date, waitlist_end_date, num_tx, REC_TX_DT, DONOR_ID, DON_TY) %>%
  filter(PERS_ID==5117188)
```

### Correct since latest waitlist end date is 1/14/2019, and transplanted 1/13/2019
```{r}
sample_collapsed_concurrent_PERS_ID %>%
  filter(PERS_ID==5117188)
```

### Check sequential to confirm
```{r}
sequential_lists %>%
  select(PERS_ID, list_type, CAN_LISTING_DT, waitlist_end_date, outcome, wait_time, REC_TX_DT, DONOR_ID, DON_TY) %>%
  filter(PERS_ID==5117188)
```



## PX_ID = 2207056 to confirm no issues with 2207056
```{r}
multiple_registrations %>%
  group_by(PERS_ID) %>%
  arrange(PERS_ID, CAN_LISTING_DT) %>% 
  mutate(list_num = row_number(),
         num_tx = n_distinct(REC_TX_DT, na.rm = TRUE)) %>%
  select(PERS_ID, list_num, list_type, CAN_LISTING_DT, min_list_date, waitlist_end_date, num_tx, REC_TX_DT, DONOR_ID, DON_TY) %>%
  filter(PERS_ID==2207056)
```

```{r}
collapsed_concurrent_registrations %>%
  select(PERS_ID, min_list_date, last_wait_date, REC_TX_DT, wait_time, outcome, DONOR_ID) %>%
  filter(PERS_ID == 2207056)
```
 


# Recombine into analytic dataset
```{r}
df_cand_kipa <- bind_rows(single_registrations %>% ungroup(), 
                          sequential_lists %>% ungroup(), 
                          collapsed_concurrent_registrations %>% ungroup())

unique_candidates <- df_cand_kipa %>% pull(PERS_ID) %>% n_distinct()
unique_listings <- df_cand_kipa %>% nrow()

rm(collapsed_concurrent_registrations, 
   multiple_registrations, single_registrations, sequential_lists,
   sample_check, sample_collapsed_concurrent_PERS_ID)
```

# Number of living donor transplants 
```{r remove_living_donors_if_requested}
num_living_donor_recips <- df_cand_kipa %>% filter(DON_TY == "L") %>% pull(PX_ID) %>% unique() %>% length()

if (living_donor_recips == FALSE) {
  df_cand_kipa <- df_cand_kipa %>%
    filter(DON_TY != "L")
}
```



# Add EPTS variables
## Dialysis time
```{r}
df_cand_kipa <- df_cand_kipa %>% 
  mutate(
    dialysis_time_at_list = case_when(
      is.na(CAN_DIAL_DT) | as.Date(min_list_date) < as.Date(CAN_DIAL_DT) ~ 0,
      TRUE ~ as.numeric(difftime(min_list_date, CAN_DIAL_DT, units = 'weeks')) / 52.25),
    preemptive_listing = case_when(
      dialysis_time_at_list == 0 ~ 1,
      TRUE ~ 0),
    dialysis = case_when(
      preemptive_listing == 0 ~ 1,
      TRUE ~ 0),
    living_donor = ifelse(outcome == "LDKT", 1, 0)
    )


df_cand_kipa$dialysis_time_at_transplant <- as.numeric(difftime(df_cand_kipa$REC_TX_DT, df_cand_kipa$CAN_DIAL_DT, units = 'weeks')) / 52.25
df_cand_kipa$dialysis_time_at_transplant[is.na(df_cand_kipa$CAN_DIAL_DT)] <- 0
df_cand_kipa$dialysis_time_at_transplant[is.na(df_cand_kipa$REC_TX_DT)] <- NA
df_cand_kipa$preemptive_transplant <- ifelse(df_cand_kipa$dialysis_time_at_transplant == 0, 1, 0)
```


## Dialysis time at listing (in years) check
```{r}
df_cand_kipa %>%
  select(min_list_date, CAN_DIAL_DT, dialysis_time_at_list, preemptive_listing, dialysis)
```

## Dialysis time at transplant check
```{r}
df_cand_kipa %>%
  select(CAN_DIAL_DT, REC_TX_DT, dialysis_time_at_transplant, preemptive_transplant, living_donor, DON_TY) %>%
  filter(is.na(REC_TX_DT) == FALSE)
```


## DM categories
```{r}
df_cand_kipa <- df_cand_kipa %>%
  mutate(
    diabetes_cat = case_when(
      CAN_DIAB_TY == '2' | CAN_DIAB_TY == '3' | CAN_DIAB_TY == '4' | CAN_DIAB_TY == '5' ~ 1,
      TRUE ~ 0))
```



# Calculate Raw EPTS at listing
```{r}
df_cand_kipa <- df_cand_kipa %>% 
  mutate(raw_epts = 
           0.047*pmax(CAN_AGE_AT_LISTING - 25, 0) - 
           0.015*(diabetes_cat==1)*pmax(CAN_AGE_AT_LISTING - 25, 0) +
           0.398*CAN_PREV_TX - 0.237*(diabetes_cat==1)*CAN_PREV_TX +
           0.315*log(dialysis_time_at_list + 1) - 0.099*(diabetes_cat==1)*log(dialysis_time_at_list + 1) +
           0.130*(dialysis_time_at_list == 0) - 0.348*(dialysis_time_at_list == 0) +  
           1.262*(diabetes_cat==1))
```


# NA EPTS values
```{r}
df_cand_kipa %>%
  filter(is.na(raw_epts)) %>%
  select(PERS_ID, CAN_AGE_AT_LISTING, diabetes_cat, CAN_PREV_TX, 
         dialysis_time_at_list, CAN_DIAL_DT, CAN_LISTING_DT, PX_ID, min_list_date)
```



# EPTS mapping to percentiles
```{r}
df_cand_kipa <- df_cand_kipa %>%
  mutate(percentile_epts = case_when(
           raw_epts <= 0.01842385502984 ~ 0,
           raw_epts <= 0.23587063655031 ~ 1,
           raw_epts <= 0.41885215605749 ~ 2,
           raw_epts <= 0.52800000000000 ~ 3,
           raw_epts <= 0.62852561015766 ~ 4,
           
           raw_epts <= 0.71352703627652 ~ 5,
           raw_epts <= 0.79220296573099 ~ 6,
           raw_epts <= 0.86714433496842 ~ 7,
           raw_epts <= 0.93047980835044 ~ 8,
           raw_epts <= 0.99391854893908 ~ 9,
           
           raw_epts <= 1.05598976270380 ~ 10,
           raw_epts <= 1.11434794060852 ~ 11,
           raw_epts <= 1.17045574165807 ~ 12,
           raw_epts <= 1.22030049020253 ~ 13,
           raw_epts <= 1.27212500074301 ~ 14,
           
           raw_epts <= 1.31913766912690 ~ 15,
           raw_epts <= 1.36438535249829 ~ 16,
           raw_epts <= 1.41168240930869 ~ 17,
           raw_epts <= 1.45433196440794 ~ 18,
           raw_epts <= 1.49473716632444 ~ 19,
           
           raw_epts <= 1.53514236824093 ~ 20,
           raw_epts <= 1.57300507228037 ~ 21,
           raw_epts <= 1.60810335386721 ~ 22,
           raw_epts <= 1.64297604380561 ~ 23,
           raw_epts <= 1.67354110859117 ~ 24,
           
           raw_epts <= 1.70255373032170 ~ 25,
           raw_epts <= 1.73150650239562 ~ 26,
           raw_epts <= 1.75869376757638 ~ 27,
           raw_epts <= 1.78477960301164 ~ 28,
           raw_epts <= 1.81180219028063 ~ 29,
           
           raw_epts <= 1.83683495289066 ~ 30,
           raw_epts <= 1.86060232717317 ~ 31,
           raw_epts <= 1.88375164234518 ~ 32,
           raw_epts <= 1.90727234572108 ~ 33,
           raw_epts <= 1.92941067761807 ~ 34,
           
           raw_epts <= 1.95149964313427 ~ 35,
           raw_epts <= 1.97370841889117 ~ 36,
           raw_epts <= 1.99414162335253 ~ 37,
           raw_epts <= 2.01486542476192 ~ 38,
           raw_epts <= 2.03323308735222 ~ 39,
           
           raw_epts <= 2.05397604380561 ~ 40,
           raw_epts <= 2.07327619854998 ~ 41,
           raw_epts <= 2.09202182261888 ~ 42,
           raw_epts <= 2.11055715263518 ~ 43,
           raw_epts <= 2.12710952613370 ~ 44,
           
           raw_epts <= 2.14469472963723 ~ 45,
           raw_epts <= 2.16206639288159 ~ 46,
           raw_epts <= 2.17895804742641 ~ 47,
           raw_epts <= 2.19533620671360 ~ 48,
           raw_epts <= 2.21202975619422 ~ 49,
           
           raw_epts <= 2.22830663928816 ~ 50,
           raw_epts <= 2.24486516084873 ~ 51,
           raw_epts <= 2.25934770704997 ~ 52,
           raw_epts <= 2.27457912196537 ~ 53,
           raw_epts <= 2.29007370896616 ~ 54,
           
           raw_epts <= 2.30555527524319 ~ 55,
           raw_epts <= 2.32162833675565 ~ 56,
           raw_epts <= 2.33676297950084 ~ 57,
           raw_epts <= 2.35264238040105 ~ 58,
           raw_epts <= 2.36804654346338 ~ 59,
           
           raw_epts <= 2.38288012607080 ~ 60,
           raw_epts <= 2.39744314980323 ~ 61,
           raw_epts <= 2.41176454483231 ~ 62,
           raw_epts <= 2.42661080994812 ~ 63,
           raw_epts <= 2.44143904753026 ~ 64,
           
           raw_epts <= 2.45616599664439 ~ 65,
           raw_epts <= 2.47046406570842 ~ 66,
           raw_epts <= 2.48541673804140 ~ 67,
           raw_epts <= 2.50082593655723 ~ 68,
           raw_epts <= 2.51575906913073 ~ 69,
           
           raw_epts <= 2.53179041411307 ~ 70,
           raw_epts <= 2.54847145496397 ~ 71,
           raw_epts <= 2.56450211494358 ~ 72,
           raw_epts <= 2.58056343072585 ~ 73,
           raw_epts <= 2.59622536132146 ~ 74,
           
           raw_epts <= 2.61220575225961 ~ 75,
           raw_epts <= 2.62980229727812 ~ 76,
           raw_epts <= 2.64633742797854 ~ 77,
           raw_epts <= 2.66270836678796 ~ 78,
           raw_epts <= 2.68071428443369 ~ 79,
           
           raw_epts <= 2.69965721597721 ~ 80,
           raw_epts <= 2.71924640657084 ~ 81,
           raw_epts <= 2.73853621699103 ~ 82,
           raw_epts <= 2.75778147760295 ~ 83,
           raw_epts <= 2.77728756159785 ~ 84,
           
           raw_epts <= 2.79719224080293 ~ 85,
           raw_epts <= 2.81884538728167 ~ 86,
           raw_epts <= 2.83956718767238 ~ 87,
           raw_epts <= 2.86117343797221 ~ 88,
           raw_epts <= 2.88458357309819 ~ 89,
           
           raw_epts <= 2.90679669457193 ~ 90,
           raw_epts <= 2.93187604974332 ~ 91,
           raw_epts <= 2.95563732458239 ~ 92,
           raw_epts <= 2.98192663471240 ~ 93,
           raw_epts <= 3.01041956645050 ~ 94,
           
           raw_epts <= 3.04211557588171 ~ 95,
           raw_epts <= 3.07653689699372 ~ 96,
           raw_epts <= 3.11838513945413 ~ 97,
           raw_epts <= 3.17159971150431 ~ 98,
           raw_epts <= 3.24849418211424 ~ 99,
           TRUE ~ 100)) 
  
## Extra line to remove missings
df_cand_kipa$percentile_epts[is.na(df_cand_kipa$raw_epts)] <- NA

df_cand_kipa <- df_cand_kipa %>%
  mutate(top_percentile_epts = case_when(
    percentile_epts <= 20 & !is.na(percentile_epts) ~ '1', 
    TRUE ~ '0')) 


df_cand_kipa$top_percentile_epts = factor(df_cand_kipa$top_percentile_epts, levels=c(0,1))
```


```{r}
df_cand_kipa %>%
  select(CAN_AGE_AT_LISTING, CAN_PREV_TX, dialysis_time_at_list, preemptive_listing, diabetes_cat, raw_epts, percentile_epts, top_percentile_epts)
```

## Age, diabetes, and preemptive vs. EPTS percentile check
```{r}
df_cand_kipa %>%
  ggplot(aes(x = CAN_AGE_AT_LISTING, y = percentile_epts, color = factor(preemptive_listing), shape = factor(diabetes_cat)))+
  geom_point()
```


## Examine cases where EPTS is 100
## The lower cutoff for 100% EPTS is roughly 3.248
```{r}
df_cand_kipa %>%
  filter(percentile_epts == 100) %>%
  select(PX_ID, raw_epts, CAN_AGE_AT_LISTING, CAN_RACE, diabetes_cat, CAN_AGE_AT_LISTING, dialysis_time_at_list, CAN_PREV_TX)
```



# Demographic variables
```{r}
df_cand_kipa <- df_cand_kipa %>%
  mutate(race = case_when(
    CAN_RACE == '8' ~ 'White',
    CAN_RACE == '16' ~ 'Black_AA',
    CAN_RACE == '32' ~ 'American_Indian',
    CAN_RACE == '64' ~ 'Asian',
    CAN_RACE == '128' ~ 'Pacific_Islander',
    CAN_RACE == '256' ~ 'Arab_Middle_East',
    CAN_RACE == '512' ~ 'Indian_SubContinent',
    CAN_RACE == '1024' ~ 'Unknown',
    CAN_RACE == '2000' ~ 'Hispanic',
    TRUE ~ 'Other')) %>%
  
  mutate(sex = case_when(
    CAN_GENDER == 'M' ~ 'Male',
    CAN_GENDER == 'F' ~ 'Female'),
    
    diabetes = factor(diabetes_cat, levels=c(0,1)),
    dialysis = factor(dialysis, levels=c(0,1)),
    preemptive_transplant = factor(preemptive_transplant, levels=c(0,1)),
    preemptive_listing = factor(preemptive_listing, levels=c(0,1)),
    previous_TX = factor(CAN_PREV_TX, levels=c(0,1)),
    age = as.numeric(CAN_AGE_AT_LISTING)
  ) %>%
  
  mutate(
    age_group = case_when(
      age <= 24 ~ '18-24',
      age > 24 & age <= 29 ~ '25-29',
      age > 29 & age <= 39 ~ '30-39',
      age > 39 & age <= 49 ~ '40-49',
      age > 49 & age <= 59 ~ '50-59',
      age > 59 & age <= 69 ~ '60-69',
      TRUE ~ '>70')) %>%
  
  select(-c(CAN_RACE, CAN_PREV_TX, CAN_AGE_AT_LISTING, diabetes_cat))

df_cand_kipa$age_group <- factor(df_cand_kipa$age_group, levels=c('18-24', '25-29', '30-39',
                                                                  '40-49', '50-59', '60-69', '>70')) 
df_cand_kipa$race <- factor(df_cand_kipa$race, levels=c('White', 'American_Indian', 'Arab_Middle_East',
                                                        'Asian', 'Black_AA', 'Hispanic', 'Indian_SubContinent',
                                                        'Other', 'Pacific_Islander', 'Unknown'))
```


# KDPI scores
```{r}
df_kdpi <- read.csv('./fulldata.csv')
df_kdpi <- df_kdpi[ ,c('DONOR_ID', 'kdpi')]
df_kdpi <- df_kdpi[!duplicated(df_kdpi), ]

df_cand_kipa %>%
  filter(DON_TY == "C")

df_cand_kipa <- merge(df_cand_kipa, df_kdpi, by = 'DONOR_ID', all.x = T, all.y = F)
```


# START HERE- missing KDPI for values for many DDKT transplants, need to check source file
```{r}
summary(df_cand_kipa %>% filter(DON_TY == "C") %>% pull(kdpi))
```


```{r}
df_cand_kipa_all %>%
  filter(PX_ID == 1020156) %>%
  select(PX_ID, CAN_LISTING_DT, REC_TX_DT, DON_TY, DONOR_ID)
```

```{r}
df_cand_kipa %>%
  filter(PX_ID == 1020156) %>%
  select(PX_ID, CAN_LISTING_DT, REC_TX_DT, DON_TY, kdpi)

unique_candidates
```


## Label top 20% KDPI category and check outcomes
```{r}
# df_cand_kipa <- df_cand_kipa %>%
#   left_join(df_kdpi) %>%
#   mutate(outcome = case_when(
#     kdpi <= 20 ~ "top 20% KDPI DDKT",
#     is.na(kdpi) == FALSE ~ ">20% KDPI DDKT",
#     TRUE ~ outcome
#   ))
# 
# df_cand_kipa %>% 
#   select(PERS_ID, outcome, kdpi, DONOR_ID) %>%
#   group_by(outcome) %>%
#   summarise(min_kdpi = min(kdpi),
#             max_kdpi = max(kdpi),
#             num_outcomes = n())
```

## Select final variables
```{r}
##Add after KDPI fixed
```


# Exclude listings missing key variables
```{r exclusions}
df_cand_kipa <- df_cand_kipa %>%
  drop_na(diabetes)

n_no_dm <- unique_listings - (df_cand_kipa$PX_ID %>% unique() %>% length())
  
df_cand_kipa <- df_cand_kipa %>%
  drop_na(age)

n_no_age <- unique_listings - n_no_dm - (df_cand_kipa$PX_ID %>% unique() %>% length())

df_cand_kipa <- df_cand_kipa %>%
  drop_na(previous_TX)

n_no_prev_tx <- unique_listings - (n_no_dm + n_no_age + (df_cand_kipa$PX_ID %>% unique() %>% length()))


df_cand_kipa <- df_cand_kipa %>%
  drop_na(race)

n_no_race <- unique_listings - (n_no_dm + n_no_age + n_no_prev_tx + (df_cand_kipa$PX_ID %>% unique() %>% length()))

```

# Write file out
```{r write_out}
if (living_donor_recips == TRUE) {
  f_name <- "candidates_w_ldkt_recips.Rdata"
} else {
  f_name <- "candidates_NO_ldkt_recips.Rdata"
}

save(df_cand_kipa, file = f_name)
```



