---
title: "Analysis of Candidates Data, with Living Donors"
output: 
  html_document:
    toc: true
---

<!-- All adult deceased donor candidates/Table 1 (Line 39) -->
<!-- Average EPTS by age group and race, deceased donors/Figure 1A (Line 66) -->
<!-- Average EPTS by age group and race, deceased donors, table (Line 109)  -->
<!-- Proportion of candidates with EPTS < 20 by age group and race, deceased donors/Figure 1B (Line 140)  -->
<!-- Logistic regression for EPTS, includes living donors (Line 190) -->
<!-- Logistic regression for KDPI, includes living donors (Line 204)  -->
<!-- Survival plots for competing risks/Figure 2 (Line 228) -->
<!-- Overall results for top 20% EPTS only (Line 326) -->
<!-- Adjusted results for age (Line 353) -->


```{r, include=F, warning=F, echo=F, message=F, comment=NA}
rm(list=ls()); gc()
setwd('C:/Users/kevinz94/Desktop/EPTS_Score')
if (!require('pacman')) {install.packages('pacman')}
library(pacman)
pacman::p_load(tidyverse, Hmisc, survival, survminer, ggplot2, gtsummary, knitr, kableExtra, 
               tidycmprsk, ggsurvfit, cowplot)
knitr::opts_chunk$set(echo = FALSE)

load('./candidates_livingdonor.RData')
df_cand_kipa$race <- droplevels(df_cand_kipa$race)
df_cand_kipa$race <- factor(df_cand_kipa$race, levels=c('White', 'Black_AA', 'Hispanic', 'Asian',
                                                        'American_Indian', 'Pacific_Islander', 'Other'))
```

-----------------


### All adult deceased donor candidates (Table 1)
```{r, warning=F, echo=F, comment=NA, message=F, fig.align='center', fig.width=9}
df_cand_kipa_table1 <- df_cand_kipa %>%
  filter(living_donor == '0') %>%
  select(race, diabetes, age, sex, previous_TX, 
         preemptive_listing, dialysis_time_at_list, top_percentile_epts, top20_kdpi, kdpi)


tbl_summary(df_cand_kipa_table1[ ,c(1:7)], by = race,
            label = list(diabetes ~ 'Diabetic', 
                        age ~ 'Age at Listing',
                        sex ~ 'Gender',
                        previous_TX ~ 'Previous Transplant',
                        preemptive_listing ~ 'Pre-Emptive Listing',                        
                        dialysis_time_at_list ~ 'Years on Dialysis Pre-Listing'),
            value = list(diabetes ~ '1',
                         preemptive_listing ~ '1',
                         previous_TX ~ '1'),
            statistic = list(all_continuous() ~ '{mean} ({sd})')) %>% 
  add_overall() %>%
  bold_labels() %>%
  modify_caption("**Table 1. All adult deceased donor candidates from 2015 to 2020**") 
```

---------------------


### Average EPTS by age group and race, deceased donors (Figure 1A)
```{r, warning=F, echo=F, comment=NA, message=F, fig.align='center'}
df_cand_kipa_fig1 <- 
  df_cand_kipa %>% filter(living_donor == '0') %>%
     group_by(race, age_group) %>%
     summarise(median = median(percentile_epts),
               low_iqr = quantile(percentile_epts, probs=0.25),
               up_iqr = quantile(percentile_epts, probs=0.75),
               mean_t = mean(percentile_epts),
               sd_t = sd(percentile_epts),
               n_candidates = n()) %>%
  
  mutate(se = sd_t / sqrt(n_candidates-1),
         low_ci_mean = mean_t - 1.96*se,
         up_ci_mean = mean_t + 1.96*se) %>%
  
  filter(race == 'White' | race == 'Black_AA' | race == 'Hispanic' | race == 'Asian') %>%
  
  mutate(race = factor(race, levels=c('Black_AA', 'Hispanic', 'White', 'Asian')))


ggplot(df_cand_kipa_fig1, 
       aes(x = age_group, y = mean_t, 
           color = race, group = race)) +
  geom_point(size=1.25) + geom_line(size=1.05) +
  geom_errorbar(aes(ymin=low_ci_mean, ymax=up_ci_mean), width=0.5, size=0.5) + 
  theme_bw() + 
  xlab('Age Group (years)') + ylab('Average EPTS Percentile') + labs(col='Race') +
  scale_color_manual(values = c('firebrick', '#009E73', '#56B4E9', '#CC79A7'), 
                     labels = c('Black or AA', 'Hispanic/Latino', 'White or Caucasian', 'Asian')) +
  theme(
    panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
    axis.title.x = element_text(size = 12, margin = margin(t=5)), 
    axis.text.y = element_text(size = 12, margin = margin(r=5)),
    plot.title = element_text(face = 'bold', size = 14, hjust=0.5)) 


rm(df_cand_kipa_fig1)
```

-----------------


### Average EPTS by age group and race, deceased donors, table
```{r, warning=F, echo=F, comment=NA, message=F, fig.align='center'}
df_cand_kipa_table2 <- df_cand_kipa %>% filter(living_donor == '0') %>%
  group_by(race, age_group) %>%
  summarise(avg_percentile_epts = round(mean(percentile_epts), 2),
            lower = round(quantile(percentile_epts, 0.25), 2),
            upper = round(quantile(percentile_epts, 0.75), 2)) %>%
  
  mutate(result = paste0(avg_percentile_epts, ' (', lower, ' - ', upper, ')')) %>%
  
  select(race, age_group, result) %>%
  
  pivot_wider(names_from = race, values_from = result)


colnames(df_cand_kipa_table2) <- c('Age Group', 'White or Caucasian', 'Black or AA',
                                    'Hispanic', 'Asian', 'American_Indian',   
                                    'Pacific_Islander', 'Other')
  

kable(df_cand_kipa_table2, align=c('l', rep('c', 7))) %>%
  kable_classic() 


rm(df_cand_kipa_table2)
```

-----------------


### Proportion of candidates with EPTS < 20 by age group and race, deceased donors (Figure 1B)
```{r, warning=F, echo=F, comment=NA, message=F, fig.align='center'}
df_cand_kipa_fig2 <- df_cand_kipa %>% filter(living_donor == '0') %>%
  group_by(race, age_group) %>%
  
  filter(race == 'White' | race == 'Black_AA' | race == 'Hispanic' | race == 'Asian') %>%
  
  mutate(race = factor(race, levels=c('White', 'Black_AA', 'Hispanic', 'Asian')),
         top_percentile_epts = as.numeric(top_percentile_epts)) %>%
  
  mutate(top_percentile_epts = case_when(
    top_percentile_epts == 1 ~ 0,
    top_percentile_epts == 2 ~ 1)) %>%

  summarise(prop = as.numeric(round(100 * mean(top_percentile_epts), 2)),
            n_candidates = n()) %>%
  
  mutate(prop_decimal = prop / 100) %>%
  
  mutate(ci_lower = prop_decimal - 1.96 * 
           sqrt(prop_decimal * (1-prop_decimal) * (1/(n_candidates - 1))),
         ci_higher = prop_decimal + 1.96 * 
           sqrt(prop_decimal * (1-prop_decimal) * (1/(n_candidates - 1)))) %>%
  
  mutate(ci_lower = 100*ci_lower,
         ci_higher = 100*ci_higher)
  
  

ggplot(df_cand_kipa_fig2, 
       aes(x = age_group, y = prop, 
           color = race, group = race)) +
  geom_point(size=1.25) + geom_line(size=1.05) +
  geom_errorbar(aes(ymin=ci_lower, ymax=ci_higher), width=0.5, size=0.5) + 
  theme_bw() + 
  xlab('Age Group (years)') + ylab('Proportion in Top 20 EPTS Percentile') + labs(col='Race') +
  scale_color_manual(values = c('firebrick', '#009E73', '#56B4E9', '#CC79A7'), 
                     labels = c('White or Caucasian', 'Black or AA', 'Hispanic/Latino', 'Asian')) +
  theme(
    panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
    axis.title.x = element_text(size = 12, margin = margin(t=5)), 
    axis.text.y = element_text(size = 12, margin = margin(r=5)),
    plot.title = element_text(face = 'bold', size = 14, hjust=0.5)) 


rm(df_cand_kipa_fig2)
```

----------------------


### Logistic regression for EPTS, includes living donors
```{r, warning=F, echo=F, comment=NA, message=F, fig.align='center'}
gtsummary::tbl_regression(glm(formula = top_percentile_epts ~ age_group + race, 
                              data = df_cand_kipa, 
                              family = binomial), 
                          label = list(age_group ~ 'Age Group',
                                       race ~ 'Race'),
                          exponentiate = T) %>%
  add_significance_stars()
```

----------------------


### Logistic regression for KDPI, includes living donors
```{r, warning=F, echo=F, comment=NA, message=F, fig.align='center'}
df_cand_kipa$top20_kdpi <- factor(df_cand_kipa$top20_kdpi, levels=c(0,1))
gtsummary::tbl_regression(glm(formula = top20_kdpi ~ age_group + race, 
                              data = df_cand_kipa, 
                              family = binomial),
                          label = list(age_group ~ 'Age Group',
                                       race ~ 'Race'),
                          exponentiate=T) %>%
  add_significance_stars()
```

----------------------


### Survival plots for competing risks (Figure 2)

These levels correspond to the possible outcomes for top 20% KDPI: \newline

* 0: Censored, includes living donors
* **1: Transplant with top 20% kidney from deceased donor** (main outcome)
* 2: Transplant with non-top 20% kidney from deceased donor
* 3: Death or delisting


```{r, warning=F, echo=F, comment=NA, message=F, fig.align='center', fig.width=10, fig.height=10}
df_cand_kipa_cmp <- df_cand_kipa %>%
  
  select(PX_ID, time, race, transplant, death, top20_kdpi, removal_code, age_group, 
         top_percentile_epts, living_donor) %>%
  
  mutate(
    status_death = case_when(
      death == '1' ~ 1,
      transplant == '1' ~ 2,
      is.na(df_cand_kipa$removal_code) == T ~ 2,
      TRUE ~ 0),
    
    status_top20 = case_when(
      death == '1' ~ 3,
      top20_kdpi == '1' & living_donor == '0' ~ 1,
      top20_kdpi == '0' & living_donor == '0' ~ 2,
      is.na(df_cand_kipa$removal_code) == T ~ 3,
      TRUE ~ 0)) %>%
  
  filter(race == 'White' | race == 'Black_AA' | race == 'Hispanic' | race == 'Asian') %>%
  
  mutate(status_death = factor(status_death, levels=c(0,1,2)),
         status_top20 = factor(status_top20, levels=c(0,1,2,3))) %>%
  
  select(time, race, status_death, status_top20, age_group, top_percentile_epts)



df_cand_kipa_cmp_all_race <- df_cand_kipa %>%
  
  select(PX_ID, time, race, transplant, death, top20_kdpi, removal_code, age_group, 
         top_percentile_epts, living_donor) %>%
  
  mutate(
    status_death = case_when(
      death == '1' ~ 1,
      transplant == '1' ~ 2,
      is.na(df_cand_kipa$removal_code) == T ~ 2,
      TRUE ~ 0),
    
    status_top20 = case_when(
      death == '1' ~ 3,
      top20_kdpi == '1' & living_donor == '0' ~ 1,
      top20_kdpi == '0' & living_donor == '0' ~ 2,
      is.na(df_cand_kipa$removal_code) == T ~ 3,
      TRUE ~ 0)) %>%
  
  mutate(status_death = factor(status_death, levels=c(0,1,2)),
         status_top20 = factor(status_top20, levels=c(0,1,2,3))) %>%
  
  select(time, race, status_death, status_top20, age_group, top_percentile_epts)
```



#### Plot of Death 
```{r, warning=F, echo=F, comment=NA, message=F, fig.align='center', fig.width=10, fig.height=10}
death_cuminc <- cuminc(Surv(time / 365.25, status_death) ~ race, data = df_cand_kipa_cmp) 
ggcuminc(death_cuminc, 
         outcome = '1',
         size = 1.2) + 
  labs(x = 'Time (years)', y = 'Cumulative Incidence') +
  xlim(0, 5) +
  scale_color_manual(values = c('firebrick', '#009E73', '#56B4E9', '#CC79A7'),
                     labels = c('White or Caucasian', 'Black or AA', 'Hispanic/Latino', 'Asian')) +
  theme_bw() + add_risktable() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        axis.title.x = element_text(size = 14, margin = margin(t=5)), 
        axis.title.y = element_text(size = 14, margin = margin(r=5)),
        plot.title = element_text(face = 'bold', size = 15, hjust=0.5),
        legend.text = element_text(size = 14))  
```


#### Plot of Top 20% Transplant
```{r, warning=F, echo=F, comment=NA, message=F, fig.align='center', fig.width=10, fig.height=10}
top20_cuminc <- cuminc(Surv(time / 365.25, status_top20) ~ race, data = df_cand_kipa_cmp) 
ggcuminc(top20_cuminc, 
         outcome = '1',
         size = 1.2) + 
  labs(x = 'Time (years)', y = 'Cumulative incidence') +
  xlim(0, 5) +
  scale_color_manual(values = c('firebrick', '#009E73', '#56B4E9', '#CC79A7'),
                     labels = c('White or Caucasian', 'Black or AA', 'Hispanic/Latino', 'Asian')) +
  theme_bw() + 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        axis.title.x = element_text(size = 14, margin = margin(t=5)), 
        axis.title.y = element_text(size = 14, margin = margin(r=5)),
        axis.text.x = element_text(size = 14),
        axis.text.y = element_text(size = 14),
        legend.text = element_text(size = 14))    
```

----------------------


### Top 20% KDPI for top 20% EPTS only
```{r, warning=F, echo=F, comment=NA, message=F, fig.align='center', fig.width=10, fig.height=10}
df_cand_kipa_top_epts <- subset(df_cand_kipa_cmp, df_cand_kipa_cmp$top_percentile_epts == '1')

### Plot of Top 20% Transplant
top20_cuminc_epts <- cuminc(Surv(time / 365.25, status_top20) ~ race, data = df_cand_kipa_top_epts) 
ggcuminc(top20_cuminc, 
         outcome = '1',
         size = 1.2) + 
  labs(x = 'Time (years)', y = 'Cumulative incidence') +
  xlim(0, 5) +
  ggtitle('Cumulative Incidence of Top 20% KDPI in Top 20% EPTS Patients with Death and Delisting as Competing Risks') +
  xlim(0, 5) +
  scale_color_manual(values = c('firebrick', '#009E73', '#56B4E9', '#CC79A7'),
                     labels = c('White or Caucasian', 'Black or AA', 'Hispanic/Latino', 'Asian')) +
  theme_bw() + 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        axis.title.x = element_text(size = 14, margin = margin(t=5)), 
        axis.title.y = element_text(size = 14, margin = margin(r=5)),
        plot.title = element_text(face = 'bold', size = 10, hjust=0.5),
        legend.text = element_text(size = 14))   

```

----------------------


### Adjusted results for age
```{r, warning=F, echo=F, comment=NA, message=F, fig.align='center', fig.width=12, fig.height=12}
### Adjust age groups
crr(Surv(time / 365.25, status_top20) ~ race + age_group, data = df_cand_kipa_cmp_all_race) %>%
  tbl_regression(exp = T)
```
