---
title: "Analysis of Candidates Data, with Living Donors"
output:
  html_document:
    toc: true
    df_print: paged
authors: Kevin Zhang, Nour Asfour, and William Parker
---

<!-- All adult donor candidates/Table 1 (Line 53) -->
<!-- Average EPTS by age group and race, all donors/Figure 1A (Line 80) -->
<!-- Average EPTS by age group and race, all donors, table (Line 128)  -->
<!-- Proportion of candidates with EPTS < 20 by age group and race, all donors/Figure 1B (Line 158)  -->
<!-- Logistic regression for EPTS, includes living donors (Line 239) -->
<!-- Survival plots for competing risks/Figure 2 (Line 253) -->
<!-- Top 20% KDPI for top 20% EPTS only (Line 381) -->
<!-- Adjusted results for age (Line 417) -->


```{r, include=F, warning=F, echo=F, message=F, comment=NA}
rm(list=ls()); invisible(gc())
if (!require('pacman')) {install.packages('pacman')}
library(pacman)
pacman::p_load(tidyverse, Hmisc, survival, survminer, ggplot2, gtsummary, knitr, kableExtra, 
               tidycmprsk, ggsurvfit, cowplot)
knitr::opts_chunk$set(echo = FALSE)

load('./candidates_w_ldkt_recips.RData')
df_cand_kipa$race <- droplevels(df_cand_kipa$race)
df_cand_kipa$race <- factor(df_cand_kipa$race, levels=c('White', 'Black_AA', 'Hispanic', 'Asian',
                                                        'American_Indian', 'Pacific_Islander', 'Other'))


df_cand_kipa <- df_cand_kipa %>%
   mutate(
    top20_kdpi = case_when(
      kdpi <= 20 & !is.na(kdpi) ~ '1',
      TRUE ~ '0'),
    
    not_top20_kdpi = case_when(
      kdpi > 20 & !is.na(kdpi) ~ '1',
      TRUE ~ '0')) %>%
  
  
   mutate(
     top20_kdpi = factor(top20_kdpi, levels=c(0,1)),
     not_top20_kdpi = factor(not_top20_kdpi, levels=c(0,1)))
```

-----------------


### All adult donor candidates (Table 1)
```{r, warning=F, echo=F, comment=NA, message=F, fig.align='center', fig.width=9}
df_cand_kipa_table1 <- df_cand_kipa %>%
  select(race, diabetes, age, sex, previous_TX, 
         preemptive_listing, dialysis_time_at_list, top_percentile_epts, top20_kdpi, kdpi)


tbl_summary(df_cand_kipa_table1[ , c("race", "diabetes", "age", "sex", "previous_TX", "preemptive_listing", 
                                     "dialysis_time_at_list")], by = race,
            label = list(diabetes ~ 'Diabetic', 
                        age ~ 'Age at Listing',
                        sex ~ 'Gender',
                        previous_TX ~ 'Previous Transplant',
                        preemptive_listing ~ 'Pre-Emptive Listing',                        
                        dialysis_time_at_list ~ 'Years on Dialysis Pre-Listing'),
            value = list(diabetes ~ '1',
                         preemptive_listing ~ '1',
                         previous_TX ~ '1'),
            statistic = list(all_continuous() ~ '{mean} ({sd})')) %>% 
  add_overall() %>%
  bold_labels() %>%
  modify_caption("**Table 1. All adult donor candidates from 2015 to 2020**") 
```

---------------------


### Average EPTS by age group and race, all donors (Figure 1A)
```{r, warning=F, echo=F, comment=NA, message=F, fig.align='center'}
df_cand_kipa_fig1 <- 
  df_cand_kipa %>% 
     group_by(race, age_group) %>%
     summarise(median = median(percentile_epts),
               low_iqr = quantile(percentile_epts, probs=0.25),
               up_iqr = quantile(percentile_epts, probs=0.75),
               mean_t = mean(percentile_epts),
               sd_t = sd(percentile_epts),
               n_candidates = n()) %>%
  
  mutate(se = sd_t / sqrt(n_candidates-1),
         low_ci_mean = mean_t - 1.96*se,
         up_ci_mean = mean_t + 1.96*se) %>%
  
  filter(race == 'White' | race == 'Black_AA' | race == 'Hispanic' | race == 'Asian') %>%
  
  mutate(race = factor(race, levels=c('White', 'Black_AA', 'Hispanic', 'Asian')))


ggplot(df_cand_kipa_fig1, 
       aes(x = age_group, y = mean_t, 
           color = race, group = race)) +
  geom_point(size=1.25) + geom_line(size=1.25) +
  geom_errorbar(aes(ymin=low_ci_mean, ymax=up_ci_mean), width=0.15, size=0.25) + 
  theme_bw() + 
  xlab('Age Group (years)') + ylab('Average EPTS Score') + labs(col='Race') +
  scale_color_manual(values = c('firebrick', '#009E73', '#56B4E9', '#CC79A7'),
                     labels = c('White', 'Black or AA', 'Hispanic/Latino', 'Asian')) +
  theme(
    panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
    axis.title.x = element_text(size = 14, margin = margin(t=5)), 
    axis.text.x = element_text(size = 13, margin = margin(t=5)), 
    axis.title.y = element_text(size = 14, margin = margin(r=5)),
    axis.text.y = element_text(size = 13, margin = margin(r=5)),
    legend.title = element_text(size = 14),
    legend.text = element_text(size = 13),
    legend.position = 'bottom', legend.direction = 'horizontal',
    plot.title = element_text(face = 'bold', size = 16, hjust=0.5)) 


rm(df_cand_kipa_fig1)
```

-----------------


### Average EPTS by age group and race, all donors, table
```{r, warning=F, echo=F, comment=NA, message=F, fig.align='center'}
df_cand_kipa_table2 <- df_cand_kipa %>% 
  group_by(race, age_group) %>%
  summarise(avg_percentile_epts = round(mean(percentile_epts), 2),
            lower = round(quantile(percentile_epts, 0.25), 2),
            upper = round(quantile(percentile_epts, 0.75), 2)) %>%
  
  mutate(result = paste0(avg_percentile_epts, ' (', lower, ' - ', upper, ')')) %>%
  
  select(race, age_group, result) %>%
  
  pivot_wider(names_from = race, values_from = result)


colnames(df_cand_kipa_table2) <- c('Age Group', 'White or Caucasian', 'Black or AA',
                                    'Hispanic', 'Asian', 'American_Indian',   
                                    'Pacific_Islander', 'Other')
  

kable(df_cand_kipa_table2, align=c('l', rep('c', 7))) %>%
  kable_classic() 


rm(df_cand_kipa_table2)
```

-----------------


### Proportion of candidates with EPTS < 20 by age group and race, all donors (Figure 1B)
```{r, warning=F, echo=F, comment=NA, message=F, fig.align='center'}
df_cand_kipa_fig2 <- df_cand_kipa %>% 
  group_by(race, age_group) %>%
  
  filter(race == 'White' | race == 'Black_AA' | race == 'Hispanic' | race == 'Asian') %>%
  
  mutate(race = factor(race, levels=c('White', 'Black_AA', 'Hispanic', 'Asian')),
         top_percentile_epts = as.numeric(top_percentile_epts)) %>%
  
  mutate(top_percentile_epts = case_when(
    top_percentile_epts == 1 ~ 0,
    top_percentile_epts == 2 ~ 1)) %>%

  summarise(prop = as.numeric(round(100 * mean(top_percentile_epts), 2)),
            n_candidates = n()) %>%
  
  mutate(prop_decimal = prop / 100) %>%
  
  mutate(ci_lower = prop_decimal - 1.96 * 
           sqrt(prop_decimal * (1-prop_decimal) * (1/(n_candidates - 1))),
         ci_higher = prop_decimal + 1.96 * 
           sqrt(prop_decimal * (1-prop_decimal) * (1/(n_candidates - 1)))) %>%
  
  mutate(ci_lower = 100*ci_lower,
         ci_higher = 100*ci_higher)
  
  

ggplot(df_cand_kipa_fig2, 
       aes(x = age_group, y = prop, 
           color = race, group = race)) +
  geom_point(size=1.25) + geom_line(size=1.25) +
  geom_errorbar(aes(ymin=ci_lower, ymax=ci_higher), width=0.15, size=0.25) + 
  theme_bw() + 
  xlab('Age Group (years)') + ylab('Proportion in Top 20 EPTS Percentile') + labs(col='Race') +
  scale_color_manual(values = c('firebrick', '#009E73', '#56B4E9', '#CC79A7'),
                     labels = c('White', 'Black or AA', 'Hispanic/Latino', 'Asian')) +
 theme(
    panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
    axis.title.x = element_text(size = 14, margin = margin(t=5)), 
    axis.text.x = element_text(size = 13, margin = margin(t=5)), 
    axis.title.y = element_text(size = 14, margin = margin(r=5)),
    axis.text.y = element_text(size = 13, margin = margin(r=5)),
    legend.title = element_text(size = 14),
    legend.text = element_text(size = 13),
    legend.position = 'bottom', legend.direction = 'horizontal',
    plot.title = element_text(face = 'bold', size = 16, hjust=0.5)) 




df_cand_kipa_table3 <- df_cand_kipa %>% 
  group_by(race, age_group) %>%
  
  filter(race == 'White' | race == 'Black_AA' | race == 'Hispanic' | race == 'Asian') %>%
  
  mutate(race = factor(race, levels=c('White', 'Black_AA', 'Hispanic', 'Asian')),
         top_percentile_epts = as.numeric(top_percentile_epts)) %>%
  
  mutate(top_percentile_epts = case_when(
    top_percentile_epts == 1 ~ 0,
    top_percentile_epts == 2 ~ 1)) %>%
  
  summarise(result = round(100 * mean(top_percentile_epts), 2)) %>%
  
  select(race, age_group, result) %>% 
  
  pivot_wider(names_from = race, values_from = result)
  

kable(df_cand_kipa_table3, align=c('l', rep('c', 7))) %>%
  kable_classic()


rm(df_cand_kipa_table2, df_cand_kipa_table3)
```

----------------------


### Logistic regression for EPTS, includes living donors
```{r, warning=F, echo=F, comment=NA, message=F, fig.align='center'}
gtsummary::tbl_regression(glm(formula = top_percentile_epts ~ age_group + race, 
                              data = df_cand_kipa, 
                              family = binomial), 
                          label = list(age_group ~ 'Age Group',
                                       race ~ 'Race'),
                          exponentiate = T) %>%
  add_significance_stars()
```

----------------------


### Survival plots for competing risks (Figure 2)
These levels correspond to the possible outcomes for death: \newline

* 0: Censored
* **1: Death** (main outcome)
* 2: Transplanted or otherwise delisted


These levels correspond to the possible outcomes for top 20% KDPI: \newline

* 0: Censored, includes living donors
* **1: Transplant with top 20% kidney from deceased donor** (main outcome)
* 2: Transplant with non-top 20% kidney from deceased donor
* 3: Death or delisting


```{r, warning=F, echo=F, comment=NA, message=F, fig.align='center', fig.width=10, fig.height=10}
df_cand_kipa_cmp <- df_cand_kipa %>%
  
  select(PX_ID, wait_time, outcome, race, death, top20_kdpi, age_group, top_percentile_epts) %>%
  
  mutate(
    status_death = case_when(
      death == '1' ~ 1,
      outcome == '>20% KDPI DDKT' | outcome == 'top 20% KDPI DDKT' | outcome == 'DDKT' | outcome == 'LDKT' ~ 2,
      (outcome == 'removed/died') & death != '1' ~ 2,
      TRUE ~ 0),
    
    status_top20 = case_when(
      death == '1' ~ 3,
      outcome == 'top 20% KDPI DDKT' | (outcome == 'DDKT' & top20_kdpi == '1') ~ 1,
      outcome == '>20% KDPI DDKT' | (outcome == 'DDKT' & top20_kdpi == '0')  ~ 2,
      outcome == 'removed/died' ~ 3,
      TRUE ~ 0)) %>%
  
  filter(race == 'White' | race == 'Black_AA' | race == 'Hispanic' | race == 'Asian') %>%
  
  mutate(status_death = factor(status_death, levels=c(0,1,2)),
         status_top20 = factor(status_top20, levels=c(0,1,2,3))) %>%
  
  select(wait_time, race, status_death, status_top20, age_group, top_percentile_epts)



df_cand_kipa_cmp_all_race <- df_cand_kipa %>%
  
  select(PX_ID, wait_time, outcome, race, death, top20_kdpi, age_group, top_percentile_epts) %>%
  
  mutate(
    status_death = case_when(
      death == '1' ~ 1,
      outcome == '>20% KDPI DDKT' | outcome == 'top 20% KDPI DDKT' | outcome == 'DDKT' | outcome == 'LDKT' ~ 2,
      (outcome == 'removed/died') & death != '1' ~ 2,
      TRUE ~ 0),
    
    status_top20 = case_when(
      death == '1' ~ 3,
      outcome == 'top 20% KDPI DDKT' | (outcome == 'DDKT' & top20_kdpi == '1') ~ 1,
      outcome == '>20% KDPI DDKT' | (outcome == 'DDKT' & top20_kdpi == '0')  ~ 2,
      outcome == 'removed/died' ~ 3,
      TRUE ~ 0)) %>%
  
  mutate(status_death = factor(status_death, levels=c(0,1,2)),
         status_top20 = factor(status_top20, levels=c(0,1,2,3))) %>%
  
  select(wait_time, race, status_death, status_top20, age_group, top_percentile_epts)
```



#### Plot of Death 
```{r, warning=F, echo=F, comment=NA, message=F, fig.align='center', fig.width=10, fig.height=10}
df_cand_kipa_cmp$wait_time <- df_cand_kipa_cmp$wait_time / 365.25
df_cand_kipa_cmp$race <- factor(df_cand_kipa_cmp$race, levels = c('White', 'Black_AA', 'Hispanic', 'Asian'))

death_cuminc <- cuminc(Surv(wait_time, status_death) ~ race, data = df_cand_kipa_cmp) 
ggcuminc(death_cuminc, 
         outcome = '1',
         size = 1.2) + 
  labs(x = 'Time (years)', y = 'Cumulative Incidence') +
  xlim(0, 5) +
  scale_color_manual(values = c('firebrick', '#009E73', '#56B4E9', '#CC79A7'),
                     labels = c('White', 'Black or AA', 'Hispanic/Latino', 'Asian')) +
  theme_bw() + 
  theme(
    panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
    axis.title.x = element_text(size = 14, margin = margin(t=5)), 
    axis.text.x = element_text(size = 13, margin = margin(t=5)), 
    axis.title.y = element_text(size = 14, margin = margin(r=5)),
    axis.text.y = element_text(size = 13, margin = margin(r=5)),
    legend.title = element_text(size = 14),
    legend.text = element_text(size = 13),
    plot.title = element_text(face = 'bold', size = 16, hjust=0.5)) 
```


#### Plot of Top 20% Transplant
```{r, warning=F, echo=F, comment=NA, message=F, fig.align='center', fig.width=10, fig.height=10}
top20_cuminc <- cuminc(Surv(wait_time, status_top20) ~ race, data = df_cand_kipa_cmp) 

ggcuminc(top20_cuminc, 
         outcome = '1',
         size = 1.2) + 
  labs(x = 'Time (years)', y = 'Cumulative incidence') +
  scale_x_continuous(limits=c(0,5.46), breaks=seq(0,5,1)) +
  scale_color_manual(
    limits = c('White', 'Black_AA', 'Hispanic', 'Asian'),
    values = c('firebrick', '#009E73', '#56B4E9', '#CC79A7'),
    labels = c('White', 'Black or AA', 'Hispanic/Latino', 'Asian')) +
  theme_bw() + 
  theme(
    panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
    axis.title.x = element_text(size = 14, margin = margin(t=5)), 
    axis.text.x = element_text(size = 13, margin = margin(t=5)), 
    axis.title.y = element_text(size = 14, margin = margin(r=5)),
    axis.text.y = element_text(size = 13, margin = margin(r=5)),
    legend.title = element_text(size = 14),
    legend.text = element_text(size = 13), legend.position = 'bottom',
    plot.title = element_text(face = 'bold', size = 16, hjust=0.5)) +
  annotate("text", x = 5.45, y= 0.127, color = 'firebrick', label = '12%') +
  annotate("text", x = 5.45, y= 0.137, color = '#009E73', label = '12%') +
  annotate("text", x = 5.45, y= 0.118, color = '#56B4E9', label = '11%') +
  annotate("text", x = 5.45, y= 0.104, color = '#CC79A7', label = '10%')
```

----------------------


### Top 20% KDPI for top 20% EPTS only
```{r, warning=F, echo=F, comment=NA, message=F, fig.align='center', fig.width=10, fig.height=10}
df_cand_kipa_top_epts <- subset(df_cand_kipa_cmp, df_cand_kipa_cmp$top_percentile_epts == '1')

### Plot of Top 20% Transplant
top20_cuminc_epts <- cuminc(Surv(wait_time, status_top20) ~ race, data = df_cand_kipa_top_epts) 
ggcuminc(top20_cuminc, 
         outcome = '1',
         size = 1.2) + 
  labs(x = 'Time (years)', y = 'Cumulative incidence') +
  xlim(0, 5.2) +
  ggtitle('Cumulative Incidence of Top 20% KDPI in Top 20% EPTS Patients with Death and Delisting as Competing Risks') +
  xlim(0, 5) +
  scale_color_manual(values = c('firebrick', '#009E73', '#56B4E9', '#CC79A7'),
                     labels = c('White or Caucasian', 'Black or AA', 'Hispanic/Latino', 'Asian')) +
  theme_bw() + 
  theme(
    panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
    axis.title.x = element_text(size = 14, margin = margin(t=5)), 
    axis.text.x = element_text(size = 13, margin = margin(t=5)), 
    axis.title.y = element_text(size = 14, margin = margin(r=5)),
    axis.text.y = element_text(size = 13, margin = margin(r=5)),
    legend.title = element_text(size = 14),
    legend.text = element_text(size = 13),
    plot.title = element_text(face = 'bold', size = 16, hjust=0.5)) 
```


```{r, warning=F, echo=F, comment=NA, message=F, fig.align='center', fig.width=10, fig.height=10}
crr(Surv(wait_time, status_top20) ~ race + age_group, data = df_cand_kipa_top_epts) %>%
   tbl_regression(exp = T)
```

----------------------


### Adjusted results for age: plots
```{r, warning=F, echo=F, comment=NA, message=F, fig.align='center', fig.width=10, fig.height=10}
df_cand_kipa_cmp_all_race$wait_time <- df_cand_kipa_cmp_all_race$wait_time / 365.25
cuminc_age_adjusted <- cuminc(Surv(wait_time, status_top20) ~ race + age_group, data = df_cand_kipa_cmp_all_race)

plot_30_39 <- ggcuminc(cuminc_age_adjusted, 
         outcome = '1',
         size = 1.2) + 
  labs(x = 'Time (years)', y = 'Cumulative incidence') +
  scale_x_continuous(limits=c(0,5.46), breaks=seq(0,5,1)) + ylim(0, 0.5) +
  theme_bw() + 
  scale_color_manual(limits = c('White, 18-24', 'Black_AA, 18-24', 'Hispanic, 18-24',
                                'Asian, 18-24', 'American_Indian, 18-24', 
                                'Pacific_Islander, 18-24', 'Other, 18-24',
                                
                                'White, 25-29', 'Black_AA, 25-29', 'Hispanic, 25-29',
                                'Asian, 25-29', 'American_Indian, 25-29', 
                                'Pacific_Islander, 25-29', 'Other, 25-29',
                                
                                'White, 30-39', 'Black_AA, 30-39', 'Hispanic, 30-39',
                                'Asian, 30-39', 'American_Indian, 30-39', 
                                'Pacific_Islander, 30-39', 'Other, 30-39',
                                
                                'White, 40-49', 'Black_AA, 40-49', 'Hispanic, 40-49',
                                'Asian, 40-49', 'American_Indian, 40-49', 
                                'Pacific_Islander, 40-49', 'Other, 40-49',    
                                
                                'White, 50-59', 'Black_AA, 50-59', 'Hispanic, 50-59',
                                'Asian, 50-59', 'American_Indian, 50-59', 
                                'Pacific_Islander, 50-59', 'Other, 50-59', 
                                
                                'White, 60-69', 'Black_AA, 60-69', 'Hispanic, 60-69',
                                'Asian, 60-69', 'American_Indian, 60-69', 
                                'Pacific_Islander, 60-69', 'Other, 60-69', 
                                
                                'White, >70', 'Black_AA, >70', 'Hispanic, >70',
                                'Asian, >70', 'American_Indian, >70', 
                                'Pacific_Islander, >70', 'Other, >70'),
                     
                                values = c(rep(alpha('white', 0), 7), 
                                rep(alpha('white', 0), 7), 
                                'firebrick', '#009E73', '#56B4E9', '#CC79A7', 
                                rep(alpha('white', 0), 3),
                                rep(alpha('white', 0), 7), 
                                rep(alpha('white', 0), 7), 
                                rep(alpha('white', 0), 7), 
                                rep(alpha('white', 0), 7)), guide = 'none') +
  theme(
    panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
    axis.title.x = element_text(size = 14, margin = margin(t=5)), 
    axis.text.x = element_text(size = 13, margin = margin(t=5)), 
    axis.title.y = element_text(size = 14, margin = margin(r=5)),
    axis.text.y = element_text(size = 13, margin = margin(r=5)),
    legend.title = element_blank(),
    legend.text = element_blank(),
    plot.title = element_text(face = 'bold', size = 16, hjust=0.5)) +
  annotate("text", x = 5.36, y= 0.321, color = 'firebrick', label = '28%') +
  annotate("text", x = 5.36, y= 0.283, color = '#009E73', label = '25%') +
  annotate("text", x = 5.36, y= 0.231, color = '#56B4E9', label = '23%') +
  annotate("text", x = 4.23, y= 0.261, color = '#CC79A7', label = '25%')


plot_40_49 <- ggcuminc(cuminc_age_adjusted, 
         outcome = '1',
         size = 1.2) + 
  labs(x = 'Time (years)', y = 'Cumulative incidence') +
  xlim(0, 5.46) + ylim(0, 0.5) +
  theme_bw() + 
  scale_color_manual(limits = c('White, 18-24', 'Black_AA, 18-24', 'Hispanic, 18-24',
                                'Asian, 18-24', 'American_Indian, 18-24', 
                                'Pacific_Islander, 18-24', 'Other, 18-24',
                                
                                'White, 25-29', 'Black_AA, 25-29', 'Hispanic, 25-29',
                                'Asian, 25-29', 'American_Indian, 25-29', 
                                'Pacific_Islander, 25-29', 'Other, 25-29',
                                
                                'White, 30-39', 'Black_AA, 30-39', 'Hispanic, 30-39',
                                'Asian, 30-39', 'American_Indian, 30-39', 
                                'Pacific_Islander, 30-39', 'Other, 30-39',
                                
                                'White, 40-49', 'Black_AA, 40-49', 'Hispanic, 40-49',
                                'Asian, 40-49', 'American_Indian, 40-49', 
                                'Pacific_Islander, 40-49', 'Other, 40-49',    
                                
                                'White, 50-59', 'Black_AA, 50-59', 'Hispanic, 50-59',
                                'Asian, 50-59', 'American_Indian, 50-59', 
                                'Pacific_Islander, 50-59', 'Other, 50-59', 
                                
                                'White, 60-69', 'Black_AA, 60-69', 'Hispanic, 60-69',
                                'Asian, 60-69', 'American_Indian, 60-69', 
                                'Pacific_Islander, 60-69', 'Other, 60-69', 
                                
                                'White, >70', 'Black_AA, >70', 'Hispanic, >70',
                                'Asian, >70', 'American_Indian, >70', 
                                'Pacific_Islander, >70', 'Other, >70'),
                     
                     values = c(rep(alpha('white', 0), 7), rep(alpha('white', 0), 7), 
                                rep(alpha('white', 0), 7), 
                                'firebrick', '#009E73', '#56B4E9', '#CC79A7', rep(alpha('white', 0), 3),
                                rep(alpha('white', 0), 7), 
                                rep(alpha('white', 0), 7), rep(alpha('white', 0), 7)), guide = 'none') +
  theme(
    panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
    axis.title.x = element_text(size = 14, margin = margin(t=5)), 
    axis.text.x = element_text(size = 13, margin = margin(t=5)), 
    axis.ticks.y = element_blank(), 
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    legend.title = element_blank(),
    legend.text = element_blank(),
    plot.title = element_text(face = 'bold', size = 16, hjust=0.5)) +
  annotate("text", x = 5.40, y= 0.195, color = 'firebrick', label = '17%') +
  annotate("text", x = 5.40, y= 0.133, color = '#009E73', label = '13%') +
  annotate("text", x = 5.40, y= 0.107, color = '#56B4E9', label = '10%') +
  annotate("text", x = 5.25, y= 0.153, color = '#CC79A7', label = '13%')  


plot_50_59 <- ggcuminc(cuminc_age_adjusted, 
         outcome = '1',
         size = 1.2) + 
  labs(x = 'Time (years)', y = 'Cumulative incidence') +
  xlim(0, 5.46) + ylim(0, 0.5) +
  theme_bw() + 
  scale_color_manual(limits = c('White, 18-24', 'Black_AA, 18-24', 'Hispanic, 18-24',
                                'Asian, 18-24', 'American_Indian, 18-24', 
                                'Pacific_Islander, 18-24', 'Other, 18-24',
                                
                                'White, 25-29', 'Black_AA, 25-29', 'Hispanic, 25-29',
                                'Asian, 25-29', 'American_Indian, 25-29', 
                                'Pacific_Islander, 25-29', 'Other, 25-29',
                                
                                'White, 30-39', 'Black_AA, 30-39', 'Hispanic, 30-39',
                                'Asian, 30-39', 'American_Indian, 30-39', 
                                'Pacific_Islander, 30-39', 'Other, 30-39',
                                
                                'White, 40-49', 'Black_AA, 40-49', 'Hispanic, 40-49',
                                'Asian, 40-49', 'American_Indian, 40-49', 
                                'Pacific_Islander, 40-49', 'Other, 40-49',    
                                
                                'White, 50-59', 'Black_AA, 50-59', 'Hispanic, 50-59',
                                'Asian, 50-59', 'American_Indian, 50-59', 
                                'Pacific_Islander, 50-59', 'Other, 50-59', 
                                
                                'White, 60-69', 'Black_AA, 60-69', 'Hispanic, 60-69',
                                'Asian, 60-69', 'American_Indian, 60-69', 
                                'Pacific_Islander, 60-69', 'Other, 60-69', 
                                
                                'White, >70', 'Black_AA, >70', 'Hispanic, >70',
                                'Asian, >70', 'American_Indian, >70', 
                                'Pacific_Islander, >70', 'Other, >70'),
                                
                                values = c(rep(alpha('white', 0), 7), rep(alpha('white', 0), 7), 
                                rep(alpha('white', 0), 7), rep(alpha('white', 0), 7), 
                                'firebrick', '#009E73', '#56B4E9', '#CC79A7', rep(alpha('white', 0), 3),
                                rep(alpha('white', 0), 7), rep(alpha('white', 0), 7)), guide = 'none') +
  theme(
    panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
    axis.title.x = element_text(size = 14, margin = margin(t=5)), 
    axis.text.x = element_text(size = 13, margin = margin(t=5)), 
    axis.ticks.y = element_blank(), 
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    legend.title = element_blank(),
    legend.text = element_blank(),
    plot.title = element_text(face = 'bold', size = 16, hjust=0.5)) +
  annotate("text", x = 5.40, y= 0.111, color = 'firebrick', label = '10%') +
  annotate("text", x = 5.45, y= 0.092, color = '#009E73', label = '8%') +
  annotate("text", x = 5.45, y= 0.070, color = '#56B4E9', label = '6%') +
  annotate("text", x = 5.45, y= 0.046, color = '#CC79A7', label = '5%')  


p4 <- ggcuminc(top20_cuminc, 
         outcome = '1',
         size = 1.2) + 
  scale_x_continuous(limits=c(0,5.46), breaks=seq(0,5,1)) +
  scale_color_manual(
    limits = c('White', 'Black_AA', 'Hispanic', 'Asian'),
    values = c('firebrick', '#009E73', '#56B4E9', '#CC79A7'),
    labels = c('White', 'Black or AA', 'Hispanic/Latino', 'Asian')) +
  theme_bw() + 
  theme(legend.title = element_text(size = 14),
    legend.text = element_text(size = 13)) 

prow <- cowplot::plot_grid(plot_30_39, plot_40_49, plot_50_59, align = 'h', ncol = 3)
legend_b <- get_legend(
  p4 + guides(color = guide_legend(nrow = 1)) + theme(legend.position = 'bottom'))

cowplot::plot_grid(prow, legend_b, ncol = 1, rel_heights = c(1, .1))
```

----------------------


### Adjusted results for age: regression
```{r, warning=F, echo=F, comment=NA, message=F, fig.align='center', fig.width=12, fig.height=12}
# ### Adjust age groups
crr(Surv(wait_time / 365.25, status_top20) ~ race + age_group, data = df_cand_kipa_cmp_all_race) %>%
   tbl_regression(exp = T)
```
