---
title: "Analysis of Candidates Data"
output: 
  html_document:
    toc: true
---

<!-- All adult deceased donor candidates (Line 44) -->
<!-- Average EPTS by age group and race (Line 80) -->
<!-- Average EPTS by age group and race (Line 124)  -->
<!-- Proportion of candidates with EPTS < 20 by age group and race (Line 154)  -->
<!-- Proportion of candidates with EPTS < 20 by age group and race (Line 192) -->
<!-- -->
<!-- Average KDPI by age group and race (Line 229) -->
<!-- Average KDPI by age group and race (Line 272)  -->
<!-- Proportion of candidates with KDPI < 20 by age group and race (Line 301) -->
<!-- Proportion of candidates with KDPI < 20 by age group and race (Line 335) -->
<!-- -->
<!-- Logistic regression for EPTS (Line 370) -->
<!-- Logistic regression for KDPI (Line 382)  -->
<!-- -->
<!-- Survival plots for competing risks (Line 399) -->
<!-- Adjusted results for age (Line 463) -->


```{r, include=F, warning=F, echo=F, message=F, comment=NA}
rm(list=ls()); gc()
setwd('C:/Users/kevinz94/Desktop/EPTS_Score')
if (!require('pacman')) {install.packages('pacman')}
library(pacman)
pacman::p_load(tidyverse, Hmisc, survival, survminer, ggplot2, gtsummary, knitr, kableExtra, tidycmprsk, ggsurvfit)
knitr::opts_chunk$set(echo = FALSE)

load('./candidates.RData')
df_cand_kipa$race <- droplevels(df_cand_kipa$race)
df_cand_kipa$race <- factor(df_cand_kipa$race, levels=c('White', 'Black_AA', 'Hispanic', 'Asian',
                                                        'American_Indian', 'Pacific_Islander', 'Other'))
```


-----------------

### All adult deceased donor candidates (Table 1)
```{r, warning=F, echo=F, comment=NA, message=F, fig.align='center', fig.width=9}
df_cand_kipa_table1 <- df_cand_kipa %>%
  select(PERS_ID, race, diabetes, age, dialysis, dialysis_years_prior, time, 
         previous_TX, preemptive_listing, percentile_epts, top_percentile_epts,
         kdpi, top20_kdpi)


tbl_summary(df_cand_kipa_table1[,-1], by = race,
            label = list(diabetes ~ 'Diabetic', 
                        age ~ 'Age at Listing',
                        dialysis_years_prior ~ 'Years on Dialysis Pre-Transplant',
                        time  ~ 'Time in Days', 
                        previous_TX ~ 'Previous Transplant',
                        dialysis ~ 'On Dialysis',
                        preemptive_listing ~ 'Preemptive Listing',
                        percentile_epts ~ 'Percentile EPTS Score',
                        top_percentile_epts  ~ 'Top 20 Percentile EPTS',
                        kdpi ~ 'Percentile KDPI Score',
                        top20_kdpi ~ 'Top 20 Percentile KDPI'),
            value = list(diabetes ~ '1',
                         previous_TX ~ '1',
                         dialysis ~ '1',
                         preemptive_listing ~ '1',
                         top_percentile_epts ~ '1',
                         top20_kdpi ~ '1'),
            statistic = list(all_continuous() ~ '{mean} ({sd})')) %>% 
  add_overall() %>%
  bold_labels() %>%
  modify_caption("**Table 1. All adult deceased donor candidates from 2015 to 2020**") 

rm(df_cand_kipa_table1)
```

-----------------

### Average EPTS by age group and race (Figure 1)
```{r, warning=F, echo=F, comment=NA, message=F, fig.align='center'}
df_cand_kipa_fig1 <- 
  df_cand_kipa %>%
     group_by(race, age_group) %>%
     summarise(median = median(percentile_epts),
               low_iqr = quantile(percentile_epts, probs=0.25),
               up_iqr = quantile(percentile_epts, probs=0.75),
               mean_t = mean(percentile_epts),
               sd_t = sd(percentile_epts),
               n_candidates = n()) %>%
  
  mutate(se = sd_t / sqrt(n_candidates-1),
         low_ci_mean = mean_t - 1.96*se,
         up_ci_mean = mean_t + 1.96*se) %>%
  
  filter(race == 'White' | race == 'Black_AA' | race == 'Hispanic') %>%
  
  mutate(race = factor(race, levels=c('Black_AA', 'Hispanic', 'White')))


ggplot(df_cand_kipa_fig1, 
       aes(x = age_group, y = mean_t, 
           color = race, group = race)) +
  geom_point(size=1.25) + geom_line(size=1.05) +
  geom_errorbar(aes(ymin=low_ci_mean, ymax=up_ci_mean), width=0.5, size=0.5) + 
  theme_bw() + 
  ggtitle('Average EPTS by Age Group and Race') +
  xlab('Age Group (years)') + ylab('Average EPTS Percentile') + labs(col='Race') +
  scale_color_manual(values = c('firebrick', '#009E73', '#56B4E9'), 
                     labels = c('Black or AA', 'Hispanic/Latino', 'White or Caucasian')) +
  theme(
    panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
    axis.title.x = element_text(size = 12, margin = margin(t=5)), 
    axis.text.y = element_text(size = 12, margin = margin(r=5)),
    plot.title = element_text(face = 'bold', size = 14, hjust=0.5)) 


rm(df_cand_kipa_fig1)
```


-----------------

### Average EPTS by age group and race (Table 2)
```{r, warning=F, echo=F, comment=NA, message=F, fig.align='center'}
df_cand_kipa_table2 <- df_cand_kipa %>%
  group_by(race, age_group) %>%
  summarise(avg_percentile_epts = round(mean(percentile_epts), 2),
            lower = round(quantile(percentile_epts, 0.25), 2),
            upper = round(quantile(percentile_epts, 0.75), 2)) %>%
  
  mutate(result = paste0(avg_percentile_epts, ' (', lower, ' - ', upper, ')')) %>%
  
  select(race, age_group, result) %>%
  
  pivot_wider(names_from = race, values_from = result)


colnames(df_cand_kipa_table2) <- c('Age Group', 'White or Caucasian', 
                                    'American_Indian', 'Asian', 'Black or AA', 
                                    'Hispanic', 'Other', 'Pacific_Islander')
  

kable(df_cand_kipa_table2, align=c('l', rep('c', 7))) %>%
  kable_classic() 


rm(df_cand_kipa_table2)
```


-----------------

### Proportion of candidates with EPTS < 20 by age group and race (Figure 2)
```{r, warning=F, echo=F, comment=NA, message=F, fig.align='center'}
df_cand_kipa_fig2 <- df_cand_kipa %>%
  group_by(race, age_group) %>%
  
  filter(race == 'White' | race == 'Black_AA' | race == 'Hispanic') %>%
  
  mutate(race = factor(race, levels=c('White', 'Black_AA', 'Hispanic')),
         top_percentile_epts = as.numeric(top_percentile_epts)) %>%
  
  mutate(top_percentile_epts = case_when(
    top_percentile_epts == 1 ~ 0,
    top_percentile_epts == 2 ~ 1)) %>%

  summarise(prop = round(100 * mean(top_percentile_epts), 2))
  

ggplot(df_cand_kipa_fig2, 
       aes(x = age_group, y = prop, 
           color = race, group = race)) +
  geom_point(size=1.25) + geom_line(size=1.05) +
  theme_bw() + 
  ggtitle('Top 20 EPTS by Age Group and Selected Race') +
  xlab('Age Group (years)') + ylab('Proportion in Top 20 EPTS Percentile') + labs(col='Race') +
  scale_color_manual(values = c('firebrick', '#009E73', '#56B4E9'), 
                     labels = c('White or Caucasian', 'Black or AA', 'Hispanic/Latino')) +
  theme(
    panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
    axis.title.x = element_text(size = 12, margin = margin(t=5)), 
    axis.text.y = element_text(size = 12, margin = margin(r=5)),
    plot.title = element_text(face = 'bold', size = 14, hjust=0.5)) 


rm(df_cand_kipa_fig2)
```

-------------

### Proportion of candidates with EPTS < 20 by age group and race (Table 3)
```{r, warning=F, echo=F, comment=NA, message=F, fig.align='center'}
df_cand_kipa_table3 <- df_cand_kipa %>%
  group_by(race, age_group) %>%
  
  mutate(top_percentile_epts = as.numeric(top_percentile_epts)) %>%
  
  mutate(top_percentile_epts = case_when(
    top_percentile_epts == 1 ~ 0,
    top_percentile_epts == 2 ~ 1)) %>%
  
  summarise(result = round(100 * mean(top_percentile_epts), 2)) %>%

  select(race, age_group, result) %>% 
  
  pivot_wider(names_from = race, values_from = result)


colnames(df_cand_kipa_table3) <- c('Age Group', 'White or Caucasian', 
                                    'American_Indian', 'Asian', 'Black or AA', 
                                    'Hispanic', 'Other', 'Pacific_Islander')
  

kable(df_cand_kipa_table3, align=c('l', rep('c', 7))) %>%
  kable_classic() 


rm(df_cand_kipa_table3)
```


<!-- -->
<!-- -->
<!-- -->

----------------------

### Average KDPI by age group and race (plot)
```{r, warning=F, echo=F, comment=NA, message=F, fig.align='center'}
df_cand_kipa_kdpi1 <- 
  df_cand_kipa %>%
     group_by(race, age_group) %>%
     summarise(median = median(kdpi, na.rm=T),
               low_iqr = quantile(kdpi, probs=0.25, na.rm=T),
               up_iqr = quantile(kdpi, probs=0.75, na.rm=T),
               mean_t = mean(kdpi, na.rm=T),
               sd_t = sd(kdpi, na.rm=T),
               n_candidates = n()) %>%
  
  mutate(se = sd_t / sqrt(n_candidates-1),
         low_ci_mean = mean_t - 1.96*se,
         up_ci_mean = mean_t + 1.96*se) %>%
  
  filter(race == 'White' | race == 'Black_AA' | race == 'Hispanic') %>%
  
  mutate(race = factor(race, levels=c('Black_AA', 'Hispanic', 'White')))


ggplot(df_cand_kipa_kdpi1, 
       aes(x = age_group, y = mean_t, 
           color = race, group = race)) +
  geom_point(size=1.25) + geom_line(size=1.05) +
  geom_errorbar(aes(ymin=low_ci_mean, ymax=up_ci_mean), width=0.5, size=0.5) + 
  theme_bw() + 
  ggtitle('Average KDPI by Age Group and Race') +
  xlab('Age Group (years)') + ylab('Average KDPI Percentile') + labs(col='Race') +
  scale_color_manual(values = c('firebrick', '#009E73', '#56B4E9'), 
                     labels = c('Black or AA', 'Hispanic/Latino', 'White or Caucasian')) +
  theme(
    panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
    axis.title.x = element_text(size = 12, margin = margin(t=5)), 
    axis.text.y = element_text(size = 12, margin = margin(r=5)),
    plot.title = element_text(face = 'bold', size = 14, hjust=0.5)) 


rm(df_cand_kipa_kdpi1)
```

----------------------

### Average KDPI by age group and race (table)
```{r, warning=F, echo=F, comment=NA, message=F, fig.align='center'}
df_cand_kipa_kdpi2 <- df_cand_kipa %>%
  group_by(race, age_group) %>%
  summarise(avg_kdpi_epts = round(mean(kdpi, na.rm=T), 2),
            lower = round(quantile(kdpi, na.rm=T, 0.25), 2),
            upper = round(quantile(kdpi, na.rm=T, 0.75), 2)) %>%
  
  mutate(result = paste0(avg_kdpi_epts, ' (', lower, ' - ', upper, ')')) %>%
  
  select(race, age_group, result) %>%
  
  pivot_wider(names_from = race, values_from = result)


colnames(df_cand_kipa_kdpi2) <- c('Age Group', 'White or Caucasian', 
                                    'American_Indian', 'Asian', 'Black or AA', 
                                    'Hispanic', 'Other', 'Pacific_Islander')
  

kable(df_cand_kipa_kdpi2, align=c('l', rep('c', 7))) %>%
  kable_classic() 

rm(df_cand_kipa_kdpi2)
```


----------------------

### Proportion of candidates with KDPI < 20 by age group and race (plot)
```{r, warning=F, echo=F, comment=NA, message=F, fig.align='center'}
df_cand_kipa_kdpi3 <- df_cand_kipa %>%
  group_by(race, age_group) %>%
  
  filter(race == 'White' | race == 'Black_AA' | race == 'Hispanic') %>%
  
  mutate(race = factor(race, levels=c('White', 'Black_AA', 'Hispanic')),
         top20_kdpi = as.numeric(top20_kdpi)) %>%

  summarise(prop = round(100 * mean(top20_kdpi), 2))
  

ggplot(df_cand_kipa_kdpi3, 
       aes(x = age_group, y = prop, 
           color = race, group = race)) +
  geom_point(size=1.25) + geom_line(size=1.05) +
  theme_bw() + 
  ggtitle('Top 20 KDPI by Age Group and Selected Race') +
  xlab('Age Group (years)') + ylab('Proportion in Top 20 KDPI Percentile') + labs(col='Race') +
  scale_color_manual(values = c('firebrick', '#009E73', '#56B4E9'), 
                     labels = c('White or Caucasian', 'Black or AA', 'Hispanic/Latino')) +
  theme(
    panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
    axis.title.x = element_text(size = 12, margin = margin(t=5)), 
    axis.text.y = element_text(size = 12, margin = margin(r=5)),
    plot.title = element_text(face = 'bold', size = 14, hjust=0.5)) 


rm(df_cand_kipa_kdpi3)
```

----------------------

### Proportion of candidates with KDPI < 20 by age group and race (table)
```{r, warning=F, echo=F, comment=NA, message=F, fig.align='center'}
df_cand_kipa_kdpi4 <- df_cand_kipa %>%
  group_by(race, age_group) %>%
  
  mutate(top20_kdpi = as.numeric(top20_kdpi)) %>%
  
  summarise(result = round(100 * mean(top20_kdpi), 2)) %>%

  select(race, age_group, result) %>% 
  
  pivot_wider(names_from = race, values_from = result)


colnames(df_cand_kipa_kdpi4) <- c('Age Group', 'White or Caucasian', 
                                    'American_Indian', 'Asian', 'Black or AA', 
                                    'Hispanic', 'Other', 'Pacific_Islander')
  

kable(df_cand_kipa_kdpi4, align=c('l', rep('c', 7))) %>%
  kable_classic() 


rm(df_cand_kipa_kdpi4)
```



<!-- -->
<!-- -->
<!-- -->

----------------------

### Logistic regression for EPTS
```{r, warning=F, echo=F, comment=NA, message=F, fig.align='center'}
gtsummary::tbl_regression(glm(formula = top_percentile_epts ~ age_group + race, 
                              data = df_cand_kipa, 
                              family = binomial), 
                          label = list(age_group ~ 'Age Group',
                                       race ~ 'Race'),
                          exponentiate = T) %>%
  add_significance_stars()
```

----------------------

### Logistic regression for KDPI
```{r, warning=F, echo=F, comment=NA, message=F, fig.align='center'}
df_cand_kipa$top20_kdpi <- factor(df_cand_kipa$top20_kdpi, levels=c(0,1))
gtsummary::tbl_regression(glm(formula = top20_kdpi ~ age_group + race, 
                              data = df_cand_kipa, 
                              family = binomial),
                          label = list(age_group ~ 'Age Group',
                                       race ~ 'Race'),
                          exponentiate=T) %>%
  add_significance_stars()
```

<!-- -->
<!-- -->
<!-- -->

----------------------

### Survival plots for competing risks
```{r, warning=F, echo=F, comment=NA, message=F, fig.align='center', fig.width=15, fig.height=15}
df_cand_kipa_cmp <- df_cand_kipa %>%
  
  select(PX_ID, time, race, transplant, death, top20_kdpi, removal_code, age_group) %>%
  
  mutate(
    status_death = case_when(
      death == '1' ~ 1,
      transplant == '1' ~ 2,
      is.na(df_cand_kipa$removal_code) == T ~ 2,
      TRUE ~ 0),
    
    status_top20 = case_when(
      death == '1' ~ 2,
      top20_kdpi == '1' ~ 1,
      is.na(df_cand_kipa$removal_code) == T ~ 2,
      TRUE ~ 0)) %>%
  
  filter(race == 'White' | race == 'Black_AA' | race == 'Hispanic') %>%
  
  mutate(status_death = factor(status_death, levels=c(0,1,2)),
         status_top20 = factor(status_top20, levels=c(0,1,2))) %>%
  
  select(time, race, status_death, status_top20, age_group)


#### THESE CALCULATIONS EACH TOOK 20+ MINUTES (NON-COMPUTING CLUSTER) #### 
### Plot of Death 
death_cuminc <- cuminc(Surv(time / 365.25, status_death) ~ race, data = df_cand_kipa_cmp) 
ggcuminc(death_cuminc, 
         outcome = '1',
         size = 1.2) + 
  labs(x = 'Time (years)', y = 'Cumulative Incidence') +
  ggtitle('Cumulative Incidence of Death with Transplant and Delisting as Competing Risks') +
  scale_color_manual(values = c('firebrick', '#009E73', '#56B4E9'),
                     labels = c('White or Caucasian', 'Black or AA', 'Hispanic/Latino')) +
  theme_bw() + 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        axis.title.x = element_text(size = 12, margin = margin(t=5)), 
        axis.text.y = element_text(size = 12, margin = margin(r=5)),
        plot.title = element_text(face = 'bold', size = 14, hjust=0.5)) 


### Plot of Top 20 Transplant
top20_cuminc <- cuminc(Surv(time / 365.25, status_top20) ~ race, data = df_cand_kipa_cmp) 
ggcuminc(top20_cuminc, 
         outcome = '1',
         size = 1.2) + 
  labs(x = 'Time (years)', y = 'Cumulative incidence') +
  ggtitle('Cumulative Incidence of Top 20% KDPI with Death and Delisting as Competing Risks') +
  scale_color_manual(values = c('firebrick', '#009E73', '#56B4E9'),
                     labels = c('White or Caucasian', 'Black or AA', 'Hispanic/Latino')) +
  theme_bw() + 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        axis.title.x = element_text(size = 12, margin = margin(t=5)), 
        axis.text.y = element_text(size = 12, margin = margin(r=5)),
        plot.title = element_text(face = 'bold', size = 14, hjust=0.5))   
```

----------------------

### Adjusted results for age
```{r, warning=F, echo=F, comment=NA, message=F, fig.align='center', fig.width=9, fig.height=9}
### Plot of Top 20 Transplant, Adjusted by Age Group
top20_cuminc_age <- cuminc(Surv(time / 365.25, status_top20) ~ race + age_group, data = df_cand_kipa_cmp) 

ggcuminc(top20_cuminc_age, 
         outcome = '1',
         size = 1.2,
         linetype_aes =  T) +
  labs(x = 'Time (years)', y = 'Cumulative incidence') +
  scale_color_manual(values = c(rep('#999999', 3),
                                rep('#E69F00', 3),
                                rep('#009E73', 3),
                                rep('#F0E442', 3),
                                rep('#0072B2', 3),
                                rep('#CC79A7', 3),
                                rep('#D55E00', 3)),
                     labels = c('White or Caucasian (18-24)', 'Black or AA (18-24)', 'Hispanic/Latino (18-24)',
                                'White or Caucasian (25-29)', 'Black or AA (25-29)', 'Hispanic/Latino (25-29)',
                                'White or Caucasian (30-39)', 'Black or AA (30-39)', 'Hispanic/Latino (30-39)',
                                'White or Caucasian (40-49)', 'Black or AA (40-49)', 'Hispanic/Latino (40-49)',
                                'White or Caucasian (50-59)', 'Black or AA (50-59)', 'Hispanic/Latino (50-59)',
                                'White or Caucasian (60-69)', 'Black or AA (60-69)', 'Hispanic/Latino (60-69)',
                                'White or Caucasian (>70)', 'Black or AA (>70)', 'Hispanic/Latino (>70)')) +
  scale_linetype_manual(values = rep(c('solid', 'dotted', 'dashed'), 7),
                        labels = c('White or Caucasian (18-24)', 'Black or AA (18-24)', 'Hispanic/Latino (18-24)',
                                   'White or Caucasian (25-29)', 'Black or AA (25-29)', 'Hispanic/Latino (25-29)',
                                   'White or Caucasian (30-39)', 'Black or AA (30-39)', 'Hispanic/Latino (30-39)',
                                   'White or Caucasian (40-49)', 'Black or AA (40-49)', 'Hispanic/Latino (40-49)',
                                   'White or Caucasian (50-59)', 'Black or AA (50-59)', 'Hispanic/Latino (50-59)',
                                   'White or Caucasian (60-69)', 'Black or AA (60-69)', 'Hispanic/Latino (60-69)',
                                   'White or Caucasian (>70)', 'Black or AA (>70)', 'Hispanic/Latino (>70)')) +
  ggtitle('Cumulative Incidence of Top 20% KDPI with Death and Delisting as Competing Risks, by Age and Race') +
  theme_bw() + 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        axis.title.x = element_text(size = 12, margin = margin(t=5)), 
        axis.text.y = element_text(size = 12, margin = margin(r=5)),
        plot.title = element_text(face = 'bold', size = 14, hjust=0.5))   
```
