---
title: "Post-Transplant Analyses"
output:
  html_document:
    toc: true
    df_print: paged
authors: Kevin Zhang, Nour Asfour, and William Parker
---

<!-- All adult recipients (Line 57) -->
<!-- KM curve for death with top EPTS < 20 by race (Line 86) -->
<!-- KM curve for death with top EPTS < 20 by race, no 2015 (Line 121) -->
<!-- KM curve for death with functioning graft with top EPTS < 20 by race/Figure 3 (Line 152) -->
<!-- KM curve for death with functioning graft with top EPTS < 20 by race, no 2015 (Line 179) -->
<!-- KM curve for death with functioning graft with EPTS < 20 and top 20% KDPI by race (Line 206) -->
<!-- KM curve for death with functioning graft with EPTS < 20 and top 20% KDPI by race, no 2015 (Line 241) -->
<!-- Cumulative incidence of death with function in top EPTS, graft failure competing risk (Line 272) -->
<!-- Cumulative incidence of death with function in top EPTS, graft failure competing risk, top KDPI adjust (Line 312) -->



```{r, include=F, warning=F, echo=F, message=F, comment=NA}
rm(list=ls()); gc()
if (!require('pacman')) {install.packages('pacman')}
library(pacman)
pacman::p_load(tidyverse, Hmisc, survival, survminer, ggplot2, gtsummary, knitr, kableExtra, tidycmprsk, ggsurvfit, lubridate)
knitr::opts_chunk$set(echo = FALSE)


load('./recipients.RData')
df_recipients$race_tx <- droplevels(df_recipients$race_tx)
df_recipients$race_tx <- factor(df_recipients$race_tx, levels=c('White', 'Black_AA', 'Hispanic', 'Asian',
                                                                'American_Indian', 'Pacific_Islander', 'Other'))


df_recipients <- df_recipients %>%
   mutate(
    top20_kdpi = case_when(
      kdpi <= 20 & !is.na(kdpi) ~ '1',
      TRUE ~ '0'),
    
    not_top20_kdpi = case_when(
      kdpi > 20 & !is.na(kdpi) ~ '1',
      TRUE ~ '0')) %>%
  
  
   mutate(
     top20_kdpi = factor(top20_kdpi, levels=c(0,1)),
     not_top20_kdpi = factor(not_top20_kdpi, levels=c(0,1)))
```

-----------------


### All adult recipients
```{r, warning=F, echo=F, comment=NA, message=F, fig.align='center', fig.width=9}
df_recipients_table1 <- df_recipients %>%
  filter(living_donor == '0') %>%
  select(race_tx, diabetes_cat_tx, age_tx, sex_tx, previous_TX_KI, 
         preemptive_listing_KI, time_listing_tx, kdpi, top20_kdpi)


tbl_summary(df_recipients_table1, by = race_tx,
            label = list(diabetes_cat_tx ~ 'Diabetic', 
                        age_tx ~ 'Age at Transplant',
                        sex_tx ~ 'Gender',
                        time_listing_tx  ~ 'Waitlist Time (Listing to Transplant), Years', 
                        previous_TX_KI ~ 'Previous Transplant',
                        preemptive_listing_KI ~ 'Pre-Emptive Listing',                        
                        kdpi ~ 'KDPI Score',
                        top20_kdpi ~ 'Top20 KDPI'),
            value = list(diabetes_cat_tx ~ '1',
                         preemptive_listing_KI ~ '1',
                         previous_TX_KI ~ '1'),
            statistic = list(all_continuous() ~ '{mean} ({sd})')) %>% 
  add_overall() %>%
  bold_labels() %>%
  modify_caption("**Table 1. All adult deceased recipients from 2015 to 2020**") 
```

-----------------


### KM curve for death with top EPTS < 20 by race
```{r, warning=F, echo=F, comment=NA, message=F, fig.align='center', fig.height=9, fig.width=9}
df_recipients_top20 <- subset(df_recipients, df_recipients$top_percentile_epts == '1' & 
                                df_recipients$living_donor == '0')
df_recipients_top20 <- df_recipients_top20 %>%
  filter(race_tx == 'White' | race_tx  == 'Black_AA' | race_tx  == 'Hispanic' | race_tx  == 'Asian')
df_recipients_top20$race_tx <- factor(df_recipients_top20$race_tx, 
                                      levels=c('White', 'Black_AA', 'Hispanic', 'Asian'))


fit_top20 <- survival::survfit(Surv(time_to_death / 365.65, death) ~ race_tx, 
                               data = df_recipients_top20)

p <- ggsurvplot(fit_top20, data = df_recipients_top20, size = 1.2, censor = F,
                conf.int = F, pval = T, xlab = 'Time in Years', ylab = 'Probability', 
                legend = 'right',
                xlim = c(0,5), ylim = c(0.9, 1), pval.coord = c(4, 0.92),
                risk.table = T,        
                risk.table.col = 'strata',
                risk.table.fontsize = c(3.5),
                font.main = c(16, 'bold'),
                font.x = c(16),
                font.y = c(16),
                font.tickslab = c(14, 'bold'),
                legend.labs = c('White or Caucasian', 'Black or AA', 'Hispanic/Latino', 'Asian'),   
                ggtheme = theme_bw(),
                palette = c('firebrick', '#009E73', '#56B4E9', '#CC79A7'))
p$table <- p$table + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())
p$plot <- p$plot + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())
p
```

-----------------


### KM curve for death with top EPTS < 20 by race, no 2015
```{r, warning=F, echo=F, comment=NA, message=F, fig.align='center', fig.height=9, fig.width=9}
df_recipients_top20$year <- lubridate::year(df_recipients_top20$listing_dt_tx)
df_recipients_top20_no2015 <- subset(df_recipients_top20, df_recipients_top20$year != '2015')


fit_top20_no2015 <- survival::survfit(Surv(time_to_death / 365.25, death) ~ race_tx, 
                                      data = df_recipients_top20_no2015)

p <- ggsurvplot(fit_top20_no2015, data = df_recipients_top20_no2015, size = 1.2, censor = F,
                conf.int = F, pval = T, xlab = 'Time in Years', ylab = 'Probability', 
                legend = 'right',
                xlim = c(0,5), ylim = c(0.9, 1), pval.coord = c(4, 0.92),
                risk.table = T,        
                risk.table.col = 'strata',
                risk.table.fontsize = c(3.5),
                font.main = c(16, 'bold'),
                font.x = c(16),
                font.y = c(16),
                font.tickslab = c(14, 'bold'),
                legend.labs = c('White or Caucasian', 'Black or AA', 'Hispanic/Latino', 'Asian'),   
                ggtheme = theme_bw(),
                palette = c('firebrick', '#009E73', '#56B4E9', '#CC79A7'))
p$table <- p$table + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())
p$plot <- p$plot + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())
p
```

-----------------


### KM curve for death with functioning graft with top EPTS < 20 by race (Figure 3)
```{r, warning=F, echo=F, comment=NA, message=F, fig.align='center', fig.height=9, fig.width=9}
fit_top20 <- survival::survfit(Surv(time_to_death_function / 365.65, death_with_function) ~ race_tx, 
                               data = df_recipients_top20)

p <- ggsurvplot(fit_top20, data = df_recipients_top20, size = 1.2, censor = F,
                conf.int = T, pval = T, xlab = 'Time in Years', ylab = 'Probability', 
                legend = 'right',
                xlim = c(0,5), ylim = c(0.9, 1), pval.coord = c(4, 0.92),
                risk.table = T,        
                risk.table.col = 'strata',
                risk.table.fontsize = c(3.5),
                font.main = c(16, 'bold'),
                font.x = c(16),
                font.y = c(16),
                font.tickslab = c(14, 'bold'),
                legend.labs = c('White or Caucasian', 'Black or AA', 'Hispanic/Latino', 'Asian'),   
                ggtheme = theme_bw(),
                palette = c('firebrick', '#009E73', '#56B4E9', '#CC79A7'))
p$table <- p$table + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())
p$plot <- p$plot + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())
p
```

-----------------


### KM curve for death with functioning graft with top EPTS < 20 by race, no 2015 (Figure 3 equivalent)
```{r, warning=F, echo=F, comment=NA, message=F, fig.align='center', fig.height=9, fig.width=9}
fit_top20_no2015 <- survival::survfit(Surv(time_to_death_function / 365.25, death_with_function) ~ race_tx, 
                                      data = df_recipients_top20_no2015)

p <- ggsurvplot(fit_top20_no2015, data = df_recipients_top20_no2015, size = 1.2, censor = F,
                conf.int = F, pval = T, xlab = 'Time in Years', ylab = 'Probability', 
                legend = 'right',
                xlim = c(0,5), ylim = c(0.9, 1), pval.coord = c(4, 0.92),
                risk.table = T,        
                risk.table.col = 'strata',
                risk.table.fontsize = c(3.5),
                font.main = c(16, 'bold'),
                font.x = c(16),
                font.y = c(16),
                font.tickslab = c(14, 'bold'),
                legend.labs = c('White or Caucasian', 'Black or AA', 'Hispanic/Latino', 'Asian'),   
                ggtheme = theme_bw(),
                palette = c('firebrick', '#009E73', '#56B4E9', '#CC79A7'))
p$table <- p$table + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())
p$plot <- p$plot + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())
p
```

-----------------


### KM curve for death with functioning graft with EPTS < 20 and top 20% KDPI by race
```{r, warning=F, echo=F, comment=NA, message=F, fig.align='center', fig.height=9, fig.width=9}
df_recipients_top20_kpdi <- subset(df_recipients, df_recipients$top_percentile_epts == '1' & 
                                     df_recipients$top20_kdpi == '1' & df_recipients$living_donor == '0')
df_recipients_top20_kpdi <- df_recipients_top20_kpdi %>%
  filter(race_tx == 'White' | race_tx  == 'Black_AA' | race_tx  == 'Hispanic' | race_tx  == 'Asian')
df_recipients_top20_kpdi$race_tx <- factor(df_recipients_top20_kpdi$race_tx, 
                                           levels=c('White', 'Black_AA', 'Hispanic', 'Asian'))


fit_top20_kpdi <- survival::survfit(Surv(time_to_death_function / 365.25, death_with_function) ~ race_tx, 
                                    data = df_recipients_top20_kpdi)

p <- ggsurvplot(fit_top20_kpdi, data = df_recipients_top20_kpdi, size = 1.2, censor = F,
                conf.int = F, pval = T, xlab = 'Time in Years', ylab = 'Probability', 
                legend = 'right',
                xlim = c(0,5), ylim = c(0.9, 1), pval.coord = c(4, 0.92),
                risk.table = T,        
                risk.table.col = 'strata',
                risk.table.fontsize = c(3.5),
                font.main = c(16, 'bold'),
                font.x = c(16),
                font.y = c(16),
                font.tickslab = c(14, 'bold'),
                legend.labs = c('White or Caucasian', 'Black or AA', 'Hispanic/Latino', 'Asian'),   
                ggtheme = theme_bw(),
                palette = c('firebrick', '#009E73', '#56B4E9', '#CC79A7'))
p$table <- p$table + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())
p$plot <- p$plot + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())
p
```

-----------------


### KM curve for death with functioning graft with EPTS < 20 and top 20% KDPI by race, no 2015
```{r, warning=F, echo=F, comment=NA, message=F, fig.align='center', fig.height=9, fig.width=9}
df_recipients_top20_kpdi$year <-  lubridate::year(df_recipients_top20_kpdi$listing_dt_tx)
df_recipients_top20_kpdi_no2015 <- subset(df_recipients_top20_kpdi, df_recipients_top20_kpdi$year != '2015')


fit_top20_kdpi_no2015 <- survival::survfit(Surv(time_to_death_function / 365.25, death_with_function) ~ race_tx, 
                                           data = df_recipients_top20_kpdi_no2015)

p <- ggsurvplot(fit_top20_kdpi_no2015, data = df_recipients_top20_kpdi_no2015, size = 1.2, censor = F,
                conf.int = F, pval = T, xlab = 'Time in Years', ylab = 'Probability', 
                legend = 'right',
                xlim = c(0,5), ylim = c(0.9, 1), pval.coord = c(4, 0.92),
                risk.table = T,        
                risk.table.col = 'strata',
                risk.table.fontsize = c(3.5),
                font.main = c(16, 'bold'),
                font.x = c(16),
                font.y = c(16),
                font.tickslab = c(14, 'bold'),
                legend.labs = c('White or Caucasian', 'Black or AA', 'Hispanic/Latino', 'Asian'),   
                ggtheme = theme_bw(),
                palette = c('firebrick', '#009E73', '#56B4E9', '#CC79A7'))
p$table <- p$table + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())
p$plot <- p$plot + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())
p
```

-----------------


### Cumulative incidence of death with function in top EPTS, graft failure competing risk
```{r, warning=F, echo=F, comment=NA, message=F, fig.align='center', fig.width=9, fig.height=9, results='hide'}
df_recipients_top20 <- df_recipients_top20 %>%
  
  mutate(status_death_with_function = case_when(
    death_with_function == '1' ~ 1,
    graft_failure == '1' ~ 2,
    TRUE ~ 0),
    
    time_death_with_function = case_when(
      status_death_with_function == '1' ~ time_to_death_function,
      status_death_with_function == '2' ~ time_graft_failure,
      TRUE ~ time_fu)) %>%
  
  mutate(status_death_with_function = factor(status_death_with_function, levels = c(0,1,2)))
  

death_function_cuminc <- cuminc(Surv(time_death_with_function / 365.25, status_death_with_function) ~ race_tx, 
                                data = df_recipients_top20) 

ggcuminc(death_function_cuminc, 
         outcome = '1', size = 1.2) + 
  labs(x = 'Time (years)', y = 'Cumulative Incidence') +
  ggtitle('Death with Function, Graft Failure Competing Risk in Deceased Donors, Top EPTS') +
  scale_color_manual(values = c('firebrick', '#009E73', '#56B4E9', '#CC79A7')) +
  xlim(c(0,5)) + ylim(c(0,0.1)) + 
  add_pvalue(location = 'annotation', x = 4.5, y = 0.08, size = 5) +
  theme_bw() + 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        axis.title.x = element_text(size = 14, margin = margin(t=5)), 
        axis.text.x = element_text(size = 14),
        axis.title.y = element_text(size = 14, margin = margin(l=5)),
        axis.text.y = element_text(size = 14, margin = margin(r=5)),
        plot.title = element_text(face = 'bold', size = 13, hjust=0.5),
        legend.text = element_text(size = 14)) 
```

-----------------


### Cumulative incidence of death with function in top EPTS, graft failure competing risk, top KDPI adjust
```{r, warning=F, echo=F, comment=NA, message=F, fig.align='center', fig.width=9, fig.height=9, results='hide'}
death_function_cuminc_adj <- cuminc(Surv(time_death_with_function / 365.25, status_death_with_function) ~ 
                                      race_tx + top20_kdpi, 
                                    data = df_recipients_top20) 

ggcuminc(death_function_cuminc_adj, 
         outcome = '1', size = 1.2) + 
  labs(x = 'Time (years)', y = 'Cumulative Incidence') +
  ggtitle('Death with Function, Graft Failure Competing Risk in Deceased Donors, Top EPTS, Adj. Top KDPI') +
  xlim(c(0,5)) + ylim(c(0,0.1)) + 
  add_pvalue(location = 'annotation', x = 4.5, y = 0.08, size = 5) +
  scale_color_manual(
    values = c('#000000', '#E69F00', '#56B4E9', '#009E73', '#F0E442', '#0072B2', '#D55E00', '#CC79A7'),
    labels = c('White, Non Top KDPI', 'Black_AA, Non Top KDPI', 'Hispanic, Non Top KDPI', 'Asian, Non Top KDPI',
               'White, Top KDPI', 'Black_AA, Top KDPI', 'Hispanic, Top KDPI', 'Asian, Top KDPI')) +
  scale_linetype_manual(values = c(rep('solid', 4), rep('dashed', 4))) +
  theme_bw() + 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        axis.title.x = element_text(size = 14, margin = margin(t=5)), 
        axis.text.x = element_text(size = 14),
        axis.title.y = element_text(size = 14, margin = margin(l=5)),
        axis.text.y = element_text(size = 14, margin = margin(r=5)),
        plot.title = element_text(face = 'bold', size = 9, hjust=0.5),
        legend.text = element_text(size = 14)) 
```
